### 2016-03-22 paper.tex
initail commit

\documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out if you need a4paper

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{mathtools} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{tikz}
\usepackage{tabulary}
\newcommand{\hvec}{\overset{\rightharpoonup}}
\newcommand{\argmin}{\arg\!\min}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\quotes}[1]{``#1''}
\usetikzlibrary{calc,positioning, fit, arrows}
%\usepackage{biber}

\title{\LARGE \bf
ROSflight: a Lightweight, Inexpensive MAV \protect\\ Research and Development Tool
}

\author{James Jackson$^{1}$, Gary Ellingson$^{2}$, Tim McLain$^{3}$% <-this % stops a space
\thanks{$^{1}$James Jackson is a MS student in the Department of Mechanical Engineering, Brigham Young University
        {\tt\small jamesjackson@byu.edu}}%
\thanks{$^{2}$Gary Ellingson is a MS student in the Department of Mechanical Engineering, Brigham Young University
        {\tt\small gary.ellingson@byu.edu}}%
\thanks{$^{3}$Tim McLain is a professor in the Department of Mechanical Engineering, Brigham Young University
        {\tt\small mclain@byu.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

To accelerate research and development of the autonomous capabilities of micro aerial vehicles we have developed flight control framework, ROSflight, as a research tool. ROSflight makes development of autopilot code easier and more efficient by minimizing the use of embedded systems, incorporating the Robot Operating System and using off-the-shelf and open-source hardware and software. Motivation and applications for use in the research community are discussed.  Analysis of loop rate and communication bandwidth are presented as well as results from flight demonstration of two multi-rotor aircraft.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{INTRODUCTION}

In recent years there has been an explosion of research and development in autonomous capabilities for both fixed wing and multirotor micro aerial vehicles (MAVs).  The use of MAVs is begin explored across many applications, such as aerial photography, surveillance, infrastructure monitoring, and package delivery. This research interest includes large technology companies, defense contractors and hobbyists, and is being enabled by widely available, inexpensive computing power and sensors.  MAVs utilize small embedded sensors and a processor in a flight control unit (FCU). The FCU performs, at the very least, attitude estimation and control and may also accomplish high-level capabilities such as GPS waypoint following or obstacle avoidance.

Current MAV research is being performed in a variety of fields including robotics, computer science, and engineering and has included development of controllers, estimators, path planners, and novel sensing technologies.  In general, MAV research is accomplished by rapid iteration of ideas and techniques and demonstrated through rigorous simulation and testing.  Available MAV research tools, including hardware and software, are hard to use, not flexible, and often expensive, and therefore limit research productivity.

This paper outlines ROSflight, our unique research and development tool for MAVs. ROSflight utilizes a simple sensor and actuator board in tandem with a more powerful computer running the Robot Operating System (ROS). ROSflight enables the rapid development of novel autopilot algorithms by minimizing the use of embedded systems and by incorporating standardized robotics tools made publicly available through ROS.  By using this tool for MAV research and development, initial autopilot development cost and effort is reduced and research flexibility is enhanced.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{fig/hummingbird.jpg}
	\caption{Small quadrotor used for demonstrating ROSflight.}
		\label{fig:hummingbird}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{BACKGROUND}

Many FCUs have been developed for the purpose of operating MAVs and they vary widely in both price and functionality. Common FCUs include open-source projects such as the PX4~\cite{Meier2015}, Paparazzi~\cite{Gati2013}, and PixHawk~\cite{1_ardupilot.com_2016} as well as proprietary solutions such as those by MikroKopter~\cite{4_gmbh_2016}, Ascending Technologies~\cite{2_ascending_technologies_gmbh_2016}, Parrot~\cite{3_parrot.com_2016}, Lockheed Martin Procerus~\cite{kestrel.com_2016}, Cloud Cap Technologies~\cite{piccolo.com_2016} and others.  The FCU is primarily responsible for performing attitude control and stabilization, but is also used to receive measurements from embedded sensors such as IMUs, sonars, and pressure sensors. Some FCUs are built into a larger autopilot framework which includes GPS-waypoint following, path planning functionality, and vision processing~\cite{Meier2011}.  An extensive summary of autopilots can be found in~\cite{Chao2010}.

These products serve several purposes in academic research.  Some of the fully-featured autopilots are used as plug-and-play solutions for data collection and other experiments.  Others are used for developing novel guidance, navigation, control, and estimation schemes.  These latter applications often require customization of the existing code sometimes down to the level of actuation of motors and attitude stabilization.

Depending on the research goals, using fully-featured FCUs in MAV research may require extensive re-building of much of the high-level functionality.  For example, GPS-denied or GPS-degraded flight requires much of the GPS-enabled FCU functionality such as position control and waypoint following to be completely re-built to accommodate changes to fundamental navigation methods.  As a result, many research experiments have been performed with an embedded FCU performing only attitude stabilization while an onboard computer does all higher-level functionality~\cite{Engel2012,Shen2011,Haines2011,Blosch2010}.

In addition to research-oriented autopilot development, the advent of first-person-view (FPV) quadcopter racing has driven advancements in FCU technology by demanding reduced weight and faster response.  The foremost among these units is the SPRacingF3~\cite{3_seriouslypro.com_2016} (see figure~\ref{fig:spracingf3}, which runs the popular, open-source cleanflight firmware~\cite{4_cleanflight.com_2016}.  This FCU successfully reduced the weight, cost, and complexity of FCUs while also achieving up to 1000~Hz attitude control loop rates with a complementary filter performing attitude estimation.  The SPRacingF3 can commonly be purchased for approximately \$60, and weighs only 6 grams, but has performance equivalent to many of its more expensive counterparts.

With the FCU dedicated to performing only low-level sensor reading and attitude stabilization, the ROS framework~\cite{Quigley2009} can be leveraged to perform higher-level tasks on an onboard computer aboard the MAV. Besides having extensive driver support for robotic sensors and peripherals, ROS also comes with several useful development tools including a powerful simulator for testing code.  While these tools are often used in MAV research, they could be leveraged more easily using ROSflight and the SPRacingF3 FCU hardware.

\begin{figure}
	\includegraphics[width=0.5\textwidth]{fig/mikey_no_prop_guard.png}
	\caption{Large quadrotor used for demonstrating ROSflight.}
		\label{fig:mikey}
\end{figure}

\section{MOTIVATION}

While a number of commercially available FCUs currently feature high-bandwidth connections to ROS, these can be cost-prohibitive to many researchers.  These proprietary FCUs perform, particularly in attitude control, equivalently to recently developed hobby-grade FCUs that are much less expensive. Besides price, these proprietary FCUs also often embed their attitude control scheme in a way that prevents researchers from understanding the finer points of how attitude control is performed within the larger scheme of the autopilot architecture.  By using an open source FCU, this information is both available to and modifiable by the researcher, resulting in more effective architecture construction and efficient control.

Other FCUs focus more on building an all-in-one solution which strives to wrap attitude control into a larger autopilot architecture.  While these solutions offer significant capability and support, often the whole autopilot is a single embedded system.  The extensive use of embedded hardware causes many challenges which can are solved by moving development into ROS.

When developing embedded code, the processor used to produce and compile the code is different then the processor used to run the code.  As a result, embedded debugging often must happen through an in-circuit debugger such as a JTAG tool with limited functionality.  When compared with using a fully-featured integrated development environment and developing native software, debugging embedded programs can be time-consuming and difficult.

Developing the full autopilot architecture on an embedded system also makes hardware abstraction difficult.  While attempts have been made at abstraction, including by the PixHawk project~\cite{Meier2015}, embedded systems use hardware level functions that are processor specific, and can sometimes be difficult to transfer to different processors.  As processing power continues to become lighter and more affordable, expandability is an important feature to consider when developing an autopilot architecture, but is difficult to achieve when using embedded hardware.  By implementing the autopilot architecture in ROS on a supported Linux distribution the hardware becomes completely abstracted, meaning that autopilot code is neither limited to the power of a given processor nor is it tied to a specific piece of autopilot hardware.

Another benefit of the cross-compatibility of the ROS interface is that there is no discernible difference in the communication between the FCU and the onboard computer and a desktop machine.  This means that the FCU can be connected to the desktop computer during development for rapid simulation, testing and debugging, then connected to the onboard computer with less worry of the FCU communication causing problems in flight tests.

Using a full-featured Linux distribution on the onboard computer also means that development can occur directly on the onboard computer.  Many of these smaller computers, such as those used in our demonstration, offer direct access to graphical user environments, standard USB connections for a mouse and keyboard and network connections.  Given this level of interaction with the onboard computer means it can be used both as the primary development machine and for in-flight control.  In addition, having a full development environment onboard means small modifications to flight control code can be compiled directly on the machine with very little downtime and no re-flashing of FCU firmware.

\begin{figure}
\centering
  \includegraphics[width=0.2\textwidth]{fig/s-l300.jpg}
  \caption{SPRacingF3 FCU}
    \label{fig:spracingf3}
\end{figure}

Because experimental testing of MAVs incurs significant hardware setup and operating costs, MAV research and development effort is reduced by the leveraging of simulation tools.  When using embedded systems, the embedded hardware must either be simulated via software-in-the-loop (SIL) or used in a high-bandwidth connection with the simulator through hardware-in-the-loop (HIL).  Both of these systems, while effective, take time to set up and can be difficult to get working properly.  If instead development and execution of code occurs on a ROS system, SIL simulations are sufficient to test all but the embedded portion of code. This is possible because of the level of hardware abstraction of the Linux computer and significantly reduces the amount of HIL simulation required for properly testing development code.

\begin{figure}[hbtp]
  \resizebox{0.5\textwidth}{!}{
  \centering
  \input{fig/block1}
  }
  \caption{Block diagram for multirotor in standard configuration using ROSflight and a Linux computer running ROS.  The FCU accepts roll, pitch, yaw rate, and throttle commands from the computer.  It controls to these commands using a PID attitude controller and complementary filter running at speeds up to 1000~Hz.  IMU can be accessed from the FCU at 150~Hz via a low-latency USB connection, which is used to run a full-state estimator on the computer, by fusing inertial measurements with GPS, visual odometry, or motion capture.}
  \label{fig:att_block}
\end{figure}

\begin{figure}[hbtp]
  \resizebox{0.5\textwidth}{!}{
  \centering
  \input{fig/block2}
  }
  \caption{Block diagram for multirotor in rate mode using ROSflight and a Linux computer running ROS.  The FCU accepts roll rate, pitch rate, yaw rate, and throttle commands from the computer and control to these commands using gyro measurements}
  \label{fig:att_rate_block}
\end{figure}

\begin{figure}[hbtp]
  \resizebox{0.5\textwidth}{!}{
  \centering
  \input{fig/fw_block}
  }
  \caption{Block diagram for a fixed-wing MAV using ROSflight and a Linux computer running ROS.  The FCU accepts thrust and aileron, elevator, and rudder deflection commands from the computer and actuates those commands, and relays IMU, sonar and other embedded sensors at 150 back to the computer.  In this configuration, the FCU acts simply as a I/O board for the computer.  All estimation runs on the computer, in combination with position sensors such as GPS.}
  \label{fig:fw_block}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{IMPLEMENTATION}
ROSflight has been designed to leverage software developed by the open-source community and inexpensive, off-the-shelf hardware to provide a research and development tool that overcomes many of the challenges associated with other commonly used MAV FCUs. It consists of an inexpensive, embedded FCU coupled with an onboard Linux computer.  The SPRacingF3~\cite{3_seriouslypro.com_2016} is a good choice for the embedded FCU because of its very impressive performance in a small, low cost and lightweight package. The FCU's only functions are to read and filter embedded sensors, produce pule-width-modulation for actuator output, control attitude, communicate with the onboard computer and accommodate a RC safety pilot.

The onboard computer can be selected from a variety of options, including ARM-centric single-board computers like the ODROID or Raspberry Pi, or high powered computers with traditional Intel/AMD processors or NVIDIA Tegra.  The only fixed requirements are that the onboard computer has a USB 2.0 port and can run Linux and ROS.  This flexibility means that SWAP considerations can become the primary concern when choosing a processing platform and upgrading computation power becomes much more manageable.

To enable safe testing of custom software and controllers, ROSflight integrates an RC safety pilot override into the firmware.  With this override, a safety pilot operating a radio control transmitter is able to override commands sent from the onboard computer either by moving the sticks of the transmitter or flipping a transmitter switch.  During normal operation, the onboard computer has full control of the FCU through a low-latency USB connection, and the safety pilot is able to override roll, pitch, yaw rate and thrust individually with the transmitter. In the case of emergencies, flipping the switch causes the FCU to completely ignore any autopilot commands from the computer and listen only to the manual control provided by the safety pilot.

ROSflight is capable of performing both attitude estimation and control on board the FCU, taking roll, pitch, yaw rate, and throttle commands (see figure~\ref{fig:att_block}).  It can be easily configured to accept roll rate, pitch rate, yaw rate and throttle commands, or even direct actuator deflections if lower-level attitude stabilization and control is performed by the onboard control scheme (see figures~\ref{fig:att_rate_block} and \ref{fig:fw_block}). Sensor readings from the FCU are made available to the onboard computer via a USB connection at 150~Hz with less than 1~ms latency.

The FCU was modeled with several multirotor platforms in Gazebo for software-in-the-loop (SIL) simulation.  The simulator is very similar to the simulator developed in~\cite{Furrer2016}, and allows for simulation of cameras, laser scanners, IMUs, and environments, for both fixed wing and multirotor MAVs with the ROSflight system.  The interface to both the FCU and the simulator have been made identical, which makes transitioning from SIL simulation to hardware testing a seamless process.

By reducing the amount of embedded computing and leveraging the onboard computer for more flight tasks, the majority of code will be developed and run in an environment more conducive to testing and debugging, leaving only the critical real-time tasks to be performed by the embedded FCU with microsecond accuracy.

As mentioned before, the choice of onboard computer is only limited by the ability to run ROS.  However, to demonstrate compatibility with different processor architectures, the system was tested with two multirotors, carrying different onboard computers. The first was a modified AscTec hummingbird carrying an ODROID XU4 with an octa-core ARM processor and 2~GB of RAM (see figure~\ref{fig:hummingbird}). The second was a larger quadcopter carrying a Gigabyte BRIX onboard computer with an Intel i7 and 16~GB of RAM (see figure~\ref{fig:mikey}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis}

A simple analysis of bandwidth and control loop times shows that stable and accurate flight is possible for a MAV configuration using ROSflight tools.

For multirotors at the lowest level, motor commands to standard multirotor motor controllers are limited to occur at less than 400~Hz.  ROSflight, along with several other proprietary FCUs, supports special controllers with motor update rates up to 1000~Hz, which is required for multirotors with very small propellers during extremely aggressive maneuvers. Many FCUs, however, perform onboard estimation at approximately 250~Hz due to computational constraints (see table~\ref{table:mr_loop_rates}), and have demonstrated sufficiently fast response for many applications.

Since ROSflight is capable of relaying IMU messages to the onboard computer at 150~Hz, a full-state extended Kalman filter (EKF) can run on the onboard computer at 150~Hz, while the complementary filter on the FCU can execute attitude commands under feedback at the faster 1000~Hz.  This coupling of estimation and control between the onboard computer and FCU results in a fast response and accurate execution of commands and an observable full-state estimate onboard.  Drift in the FCU's internal attitude estimate is corrected with feedback from the EKF.

Fixed wing MAVs, in general, are much less dynamic than multirotors and have slower response frequencies, meaning they are less sensitive to bandwidth concerns.  This can be observed by considering the inner-loop rates of popular fixed-wing autopilots, such as those shown in table~\ref{table:fw_loop_rates}, that are much lower than those of the multirotor FCUs.  With sensor information available at 150~Hz the onboard EKF is able to achieve similar performance as these established fixed-wing autopilots.

\begin{table}
\centering
\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
\hline
       & Kestrel & MicroPilot & PixHawk & ROSFlight \\ \hline
  Loop Rate (Hz) & 100 & 30 & 250 & 150 \\ \hline
\end{tabulary}
\caption{comparison of inner control loop rates among several fixed-wing MAV FCUs}
\label{table:fw_loop_rates}
\end{table}

\begin{table}
\centering
\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
\hline
       & PX4 & PixHawk & AscTec  & ROSFlight \\ \hline
  Loop Rate (Hz) & 250 & 250 &  1000 & 1000 \\ \hline
\end{tabulary}
\caption{comparison of inner control loop rates among several multirotor MAV FCUs}
\label{table:mr_loop_rates}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RESULTS}

\begin{table}
\centering
  \begin{tabulary}{0.5\textwidth}{|C|C|C|C|}
  \hline
  message type & message size (bytes) & avg trip time (ms) & max trip time (ms) \\ \hline
  Commands     & 36                   & 0.117             & 0.2               \\ \hline
  IMU          & 22                   & 0.115             & 0.19              \\ \hline
  \end{tabulary}
  \caption{latency test results for ROSflight}
  \label{table:latency_results}
\end{table}

\begin{figure}
	\includegraphics[width=0.5\textwidth]{fig/hummingbird_att.eps}
	\includegraphics[width=0.5\textwidth]{fig/hummingbird_vel.eps}
	\caption{Position, velocity, and attitude estimates and true state plotted as a function of time for a small multirotor MAV.}
		\label{fig:hummingbird_results}
\end{figure}

Communication latency between the FCU and onboard computer was tested by sending commands to the FCU where they were registered and then requested from the FCU and returned. The send and request cycle was performed at 200~Hz.  This task required the FCU to record the incoming command, receive a request for its return, and send it, while also operating the control loops at full rate.  Because response rates may vary slightly due to the order of higher-priority routines in the FCU, 10,000 cycles were performed and the average and maximum cycle times were recorded. IMU sensor measurement latency was tested by similarly sending and immediately requesting a message of the same size as the IMU message.  Results for both test are summarized in table~\ref{table:latency_results}.  Because I$^2$C communication to the IMU is higher priority than the USB connection to the onboard computer, it can be assumed that actual IMU latency is lower than what was recorded in this test.

Besides IMU, ROSflight supports measurements from a magnetometer, a differential pressure sensor, a barometer and a sonar.  Because these sensors are used to measure much slower dynamics than accelerometer and gyroscope readings and because total communication bandwidth is limited, these sensors have been integrated and tested at 20~Hz with IMU readings at 150~Hz.  These speeds are are adjustable if required. GPS, cameras, laser scanners and other more complicated sensors used in full state estimation can be connected directly to the onboard computer with a standard USB connections.

Both multirotors shown in figures~\ref{fig:hummingbird} and~\ref{fig:mikey} were flown in a motion capture room under manual control.  An EKF was implemented on the onboard computer on both MAVs which estimated the position, velocity and attitude of the MAV as well as biases for the accelerometer $x$ and $y$ axes as well as biases for all three gyroscope axes.  A motion capture system was used to provide position measurements that were fused with IMU for full-state estimation. The estimates for velocity and attitude for the smaller quadrotor flight have been plotted alongside their respective motion capture measurements in figure~\ref{fig:hummingbird_results}. As can be seen from the results, estimates were able to track effectively throughout the flight.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{CONCLUSION}
ROSflight is a new research tool that enables rapid development and testing of autopilot code for MAVs.  It leverages hardware and software developed by the open-source community to enable accurate attitude estimation and control with a high-bandwidth, low-latency connection to ROS.  While some of these features exist in current solutions, ROSflight provides state-of-the-art performance for a fraction of the cost of these other solutions while also providing extensive flexibility in platform and onboard computer choice.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{APPENDIX}

% Appendixes should appear before the acknowledgment.

% \section*{ACKNOWLEDGMENT}

% Important people/organizations who made it possible

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{./library}
\bibliographystyle{ieeetr}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% SAVED STUFF

%new document

### 2016-03-22 paper.tex
started the outline

@@
 
-\IEEEoverridecommandlockouts                              % This command is only needed if 
+%\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                           % you want to use the \thanks command
 
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
+%\overrideIEEEmargins                                      % Needed to meet printer requirements.
 
@@
 \title{\LARGE \bf
-ROSflight: a Lightweight, Inexpensive MAV \protect\\ Research and Development Tool
+Awesome quad!!!
 }
 
-\author{James Jackson$^{1}$, Gary Ellingson$^{2}$, Tim McLain$^{3}$% <-this % stops a space
-\thanks{$^{1}$James Jackson is a MS student in the Department of Mechanical Engineering, Brigham Young University
-        {\tt\small jamesjackson@byu.edu}}%
-\thanks{$^{2}$Gary Ellingson is a MS student in the Department of Mechanical Engineering, Brigham Young University
-        {\tt\small gary.ellingson@byu.edu}}%
-\thanks{$^{3}$Tim McLain is a professor in the Department of Mechanical Engineering, Brigham Young University
-        {\tt\small mclain@byu.edu}}%
+\author{Daillin Briggs, Gary Ellingson% <-this % stops a space
+%\thanks{$^{1}$James Jackson is a MS student in the Department of Mechanical Engineering, Brigham Young University
+%        {\tt\small jamesjackson@byu.edu}}%
+%\thanks{$^{2}$Gary Ellingson is a MS student in the Department of Mechanical Engineering, Brigham Young University
+%        {\tt\small gary.ellingson@byu.edu}}%
 }
@@
 
-To accelerate research and development of the autonomous capabilities of micro aerial vehicles we have developed flight control framework, ROSflight, as a research tool.
-ROSflight makes development of autopilot code easier and more efficient by minimizing the use of embedded systems, incorporating the Robot Operating System and using off-the-shelf and open-source hardware and software.
-Motivation and applications for use in the research community are discussed.
-Analysis of loop rate and communication bandwidth are presented as well as results from flight demonstration of two multi-rotor aircraft.
+Hi this is the abstract
 
@@
 
-In recent years there has been an explosion of research and development in autonomous capabilities for both fixed wing and multirotor micro aerial vehicles (MAVs).
-The use of MAVs is begin explored across many applications, such as aerial photography, surveillance, infrastructure monitoring, and package delivery.
-This research interest includes large technology companies, defense contractors and hobbyists, and is being enabled by widely available, inexpensive computing power and sensors.
-MAVs utilize small embedded sensors and a processor in a flight control unit (FCU).
-The FCU performs, at the very least, attitude estimation and control and may also accomplish high-level capabilities such as GPS waypoint following or obstacle avoidance.
-
-Current MAV research is being performed in a variety of fields including robotics, computer science, and engineering and has included development of controllers, estimators, path planners, and novel sensing technologies.
-In general, MAV research is accomplished by rapid iteration of ideas and techniques and demonstrated through rigorous simulation and testing.
-Available MAV research tools, including hardware and software, are hard to use, not flexible, and often expensive, and therefore limit research productivity.
-
-This paper outlines ROSflight, our unique research and development tool for MAVs.
-ROSflight utilizes a simple sensor and actuator board in tandem with a more powerful computer running the Robot Operating System (ROS).
-ROSflight enables the rapid development of novel autopilot algorithms by minimizing the use of embedded systems and by incorporating standardized robotics tools made publicly available through ROS.
-By using this tool for MAV research and development, initial autopilot development cost and effort is reduced and research flexibility is enhanced.
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{fig/hummingbird.jpg}
-	\caption{Small quadrotor used for demonstrating ROSflight.}
-		\label{fig:hummingbird}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{BACKGROUND}
-
-Many FCUs have been developed for the purpose of operating MAVs and they vary widely in both price and functionality.
-Common FCUs include open-source projects such as the PX4~\cite{Meier2015}, Paparazzi~\cite{Gati2013}, and PixHawk~\cite{1_ardupilot.com_2016} as well as proprietary solutions such as those by MikroKopter~\cite{4_gmbh_2016}, Ascending Technologies~\cite{2_ascending_technologies_gmbh_2016}, Parrot~\cite{3_parrot.com_2016}, Lockheed Martin Procerus~\cite{kestrel.com_2016}, Cloud Cap Technologies~\cite{piccolo.com_2016} and others.
-The FCU is primarily responsible for performing attitude control and stabilization, but is also used to receive measurements from embedded sensors such as IMUs, sonars, and pressure sensors.
-Some FCUs are built into a larger autopilot framework which includes GPS-waypoint following, path planning functionality, and vision processing~\cite{Meier2011}.
-An extensive summary of autopilots can be found in~\cite{Chao2010}.
-
-These products serve several purposes in academic research.
-Some of the fully-featured autopilots are used as plug-and-play solutions for data collection and other experiments.
-Others are used for developing novel guidance, navigation, control, and estimation schemes.
-These latter applications often require customization of the existing code sometimes down to the level of actuation of motors and attitude stabilization.
-
-Depending on the research goals, using fully-featured FCUs in MAV research may require extensive re-building of much of the high-level functionality.
-For example, GPS-denied or GPS-degraded flight requires much of the GPS-enabled FCU functionality such as position control and waypoint following to be completely re-built to accommodate changes to fundamental navigation methods.
-As a result, many research experiments have been performed with an embedded FCU performing only attitude stabilization while an onboard computer does all higher-level functionality~\cite{Engel2012,Shen2011,Haines2011,Blosch2010}.
-
-In addition to research-oriented autopilot development, the advent of first-person-view (FPV) quadcopter racing has driven advancements in FCU technology by demanding reduced weight and faster response.
-The foremost among these units is the SPRacingF3~\cite{3_seriouslypro.com_2016} (see figure~\ref{fig:spracingf3}, which runs the popular, open-source cleanflight firmware~\cite{4_cleanflight.com_2016}.
-This FCU successfully reduced the weight, cost, and complexity of FCUs while also achieving up to 1000~Hz attitude control loop rates with a complementary filter performing attitude estimation.
-The SPRacingF3 can commonly be purchased for approximately \$60, and weighs only 6 grams, but has performance equivalent to many of its more expensive counterparts.
-
-With the FCU dedicated to performing only low-level sensor reading and attitude stabilization, the ROS framework~\cite{Quigley2009} can be leveraged to perform higher-level tasks on an onboard computer aboard the MAV.
-Besides having extensive driver support for robotic sensors and peripherals, ROS also comes with several useful development tools including a powerful simulator for testing code.
-While these tools are often used in MAV research, they could be leveraged more easily using ROSflight and the SPRacingF3 FCU hardware.
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{fig/mikey_no_prop_guard.png}
-	\caption{Large quadrotor used for demonstrating ROSflight.}
-		\label{fig:mikey}
-\end{figure}
 
@@
 
-While a number of commercially available FCUs currently feature high-bandwidth connections to ROS, these can be cost-prohibitive to many researchers.
-These proprietary FCUs perform, particularly in attitude control, equivalently to recently developed hobby-grade FCUs that are much less expensive.
-Besides price, these proprietary FCUs also often embed their attitude control scheme in a way that prevents researchers from understanding the finer points of how attitude control is performed within the larger scheme of the autopilot architecture.
-By using an open source FCU, this information is both available to and modifiable by the researcher, resulting in more effective architecture construction and efficient control.
-
-Other FCUs focus more on building an all-in-one solution which strives to wrap attitude control into a larger autopilot architecture.
-While these solutions offer significant capability and support, often the whole autopilot is a single embedded system.
-The extensive use of embedded hardware causes many challenges which can are solved by moving development into ROS.
-
-When developing embedded code, the processor used to produce and compile the code is different then the processor used to run the code.
-As a result, embedded debugging often must happen through an in-circuit debugger such as a JTAG tool with limited functionality.
-When compared with using a fully-featured integrated development environment and developing native software, debugging embedded programs can be time-consuming and difficult.
-
-Developing the full autopilot architecture on an embedded system also makes hardware abstraction difficult.
-While attempts have been made at abstraction, including by the PixHawk project~\cite{Meier2015}, embedded systems use hardware level functions that are processor specific, and can sometimes be difficult to transfer to different processors.
-As processing power continues to become lighter and more affordable, expandability is an important feature to consider when developing an autopilot architecture, but is difficult to achieve when using embedded hardware.
-By implementing the autopilot architecture in ROS on a supported Linux distribution the hardware becomes completely abstracted, meaning that autopilot code is neither limited to the power of a given processor nor is it tied to a specific piece of autopilot hardware.
-
-Another benefit of the cross-compatibility of the ROS interface is that there is no discernible difference in the communication between the FCU and the onboard computer and a desktop machine.
-This means that the FCU can be connected to the desktop computer during development for rapid simulation, testing and debugging, then connected to the onboard computer with less worry of the FCU communication causing problems in flight tests.
-
-Using a full-featured Linux distribution on the onboard computer also means that development can occur directly on the onboard computer.
-Many of these smaller computers, such as those used in our demonstration, offer direct access to graphical user environments, standard USB connections for a mouse and keyboard and network connections.
-Given this level of interaction with the onboard computer means it can be used both as the primary development machine and for in-flight control.
-In addition, having a full development environment onboard means small modifications to flight control code can be compiled directly on the machine with very little downtime and no re-flashing of FCU firmware.
-
-\begin{figure}
-\centering
-  \includegraphics[width=0.2\textwidth]{fig/s-l300.jpg}
-  \caption{SPRacingF3 FCU}
-    \label{fig:spracingf3}
-\end{figure}
-
-Because experimental testing of MAVs incurs significant hardware setup and operating costs, MAV research and development effort is reduced by the leveraging of simulation tools.
-When using embedded systems, the embedded hardware must either be simulated via software-in-the-loop (SIL) or used in a high-bandwidth connection with the simulator through hardware-in-the-loop (HIL).
-Both of these systems, while effective, take time to set up and can be difficult to get working properly.
-If instead development and execution of code occurs on a ROS system, SIL simulations are sufficient to test all but the embedded portion of code.
-This is possible because of the level of hardware abstraction of the Linux computer and significantly reduces the amount of HIL simulation required for properly testing development code.
-
-\begin{figure}[hbtp]
-  \resizebox{0.5\textwidth}{!}{
-  \centering
-  \input{fig/block1}
-  }
-  \caption{Block diagram for multirotor in standard configuration using ROSflight and a Linux computer running ROS.
-The FCU accepts roll, pitch, yaw rate, and throttle commands from the computer.
-It controls to these commands using a PID attitude controller and complementary filter running at speeds up to 1000~Hz.
-IMU can be accessed from the FCU at 150~Hz via a low-latency USB connection, which is used to run a full-state estimator on the computer, by fusing inertial measurements with GPS, visual odometry, or motion capture.}
-  \label{fig:att_block}
-\end{figure}
-
-\begin{figure}[hbtp]
-  \resizebox{0.5\textwidth}{!}{
-  \centering
-  \input{fig/block2}
-  }
-  \caption{Block diagram for multirotor in rate mode using ROSflight and a Linux computer running ROS.
-The FCU accepts roll rate, pitch rate, yaw rate, and throttle commands from the computer and control to these commands using gyro measurements}
-  \label{fig:att_rate_block}
-\end{figure}
-
-\begin{figure}[hbtp]
-  \resizebox{0.5\textwidth}{!}{
-  \centering
-  \input{fig/fw_block}
-  }
-  \caption{Block diagram for a fixed-wing MAV using ROSflight and a Linux computer running ROS.
-The FCU accepts thrust and aileron, elevator, and rudder deflection commands from the computer and actuates those commands, and relays IMU, sonar and other embedded sensors at 150 back to the computer.
-In this configuration, the FCU acts simply as a I/O board for the computer.
-All estimation runs on the computer, in combination with position sensors such as GPS.}
-  \label{fig:fw_block}
-\end{figure}
+\begin{itemize}
+	\item{Why a gas quad}
+	\item{figure of current solutions}
+\end{itemize}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{METHODS}
+analysis models
 
-\section{IMPLEMENTATION}
-ROSflight has been designed to leverage software developed by the open-source community and inexpensive, off-the-shelf hardware to provide a research and development tool that overcomes many of the challenges associated with other commonly used MAV FCUs.
-It consists of an inexpensive, embedded FCU coupled with an onboard Linux computer.
-The SPRacingF3~\cite{3_seriouslypro.com_2016} is a good choice for the embedded FCU because of its very impressive performance in a small, low cost and lightweight package.
-The FCU's only functions are to read and filter embedded sensors, produce pule-width-modulation for actuator output, control attitude, communicate with the onboard computer and accommodate a RC safety pilot.
+Empirical engine models
+aerodynamic models
+Simplifications - we are stupid
 
-The onboard computer can be selected from a variety of options, including ARM-centric single-board computers like the ODROID or Raspberry Pi, or high powered computers with traditional Intel/AMD processors or NVIDIA Tegra.
-The only fixed requirements are that the onboard computer has a USB 2.0 port and can run Linux and ROS.
-This flexibility means that SWAP considerations can become the primary concern when choosing a processing platform and upgrading computation power becomes much more manageable.
+\section{OPTIMIZATION}
+Setup the formal optimization problems
 
-To enable safe testing of custom software and controllers, ROSflight integrates an RC safety pilot override into the firmware.
-With this override, a safety pilot operating a radio control transmitter is able to override commands sent from the onboard computer either by moving the sticks of the transmitter or flipping a transmitter switch.
-During normal operation, the onboard computer has full control of the FCU through a low-latency USB connection, and the safety pilot is able to override roll, pitch, yaw rate and thrust individually with the transmitter.
-In the case of emergencies, flipping the switch causes the FCU to completely ignore any autopilot commands from the computer and listen only to the manual control provided by the safety pilot.
+What optimizers, setup, comparing matlab and python. 
+figure of Pareto fronts...
 
-ROSflight is capable of performing both attitude estimation and control on board the FCU, taking roll, pitch, yaw rate, and throttle commands (see figure~\ref{fig:att_block}).
-It can be easily configured to accept roll rate, pitch rate, yaw rate and throttle commands, or even direct actuator deflections if lower-level attitude stabilization and control is performed by the onboard control scheme (see figures~\ref{fig:att_rate_block} and \ref{fig:fw_block}).
-Sensor readings from the FCU are made available to the onboard computer via a USB connection at 150~Hz with less than 1~ms latency.
-
-The FCU was modeled with several multirotor platforms in Gazebo for software-in-the-loop (SIL) simulation.
-The simulator is very similar to the simulator developed in~\cite{Furrer2016}, and allows for simulation of cameras, laser scanners, IMUs, and environments, for both fixed wing and multirotor MAVs with the ROSflight system.
-The interface to both the FCU and the simulator have been made identical, which makes transitioning from SIL simulation to hardware testing a seamless process.
-
-By reducing the amount of embedded computing and leveraging the onboard computer for more flight tasks, the majority of code will be developed and run in an environment more conducive to testing and debugging, leaving only the critical real-time tasks to be performed by the embedded FCU with microsecond accuracy.
-
-As mentioned before, the choice of onboard computer is only limited by the ability to run ROS.
-However, to demonstrate compatibility with different processor architectures, the system was tested with two multirotors, carrying different onboard computers.
-The first was a modified AscTec hummingbird carrying an ODROID XU4 with an octa-core ARM processor and 2~GB of RAM (see figure~\ref{fig:hummingbird}).
-The second was a larger quadcopter carrying a Gigabyte BRIX onboard computer with an Intel i7 and 16~GB of RAM (see figure~\ref{fig:mikey}).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis}
-
-A simple analysis of bandwidth and control loop times shows that stable and accurate flight is possible for a MAV configuration using ROSflight tools.
-
-For multirotors at the lowest level, motor commands to standard multirotor motor controllers are limited to occur at less than 400~Hz.
-ROSflight, along with several other proprietary FCUs, supports special controllers with motor update rates up to 1000~Hz, which is required for multirotors with very small propellers during extremely aggressive maneuvers.
-Many FCUs, however, perform onboard estimation at approximately 250~Hz due to computational constraints (see table~\ref{table:mr_loop_rates}), and have demonstrated sufficiently fast response for many applications.
-
-Since ROSflight is capable of relaying IMU messages to the onboard computer at 150~Hz, a full-state extended Kalman filter (EKF) can run on the onboard computer at 150~Hz, while the complementary filter on the FCU can execute attitude commands under feedback at the faster 1000~Hz.
-This coupling of estimation and control between the onboard computer and FCU results in a fast response and accurate execution of commands and an observable full-state estimate onboard.
-Drift in the FCU's internal attitude estimate is corrected with feedback from the EKF.
-
-Fixed wing MAVs, in general, are much less dynamic than multirotors and have slower response frequencies, meaning they are less sensitive to bandwidth concerns.
-This can be observed by considering the inner-loop rates of popular fixed-wing autopilots, such as those shown in table~\ref{table:fw_loop_rates}, that are much lower than those of the multirotor FCUs.
-With sensor information available at 150~Hz the onboard EKF is able to achieve similar performance as these established fixed-wing autopilots.
-
-\begin{table}
-\centering
-\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
-\hline
-       & Kestrel & MicroPilot & PixHawk & ROSFlight \\ \hline
-  Loop Rate (Hz) & 100 & 30 & 250 & 150 \\ \hline
-\end{tabulary}
-\caption{comparison of inner control loop rates among several fixed-wing MAV FCUs}
-\label{table:fw_loop_rates}
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
-\hline
-       & PX4 & PixHawk & AscTec  & ROSFlight \\ \hline
-  Loop Rate (Hz) & 250 & 250 &  1000 & 1000 \\ \hline
-\end{tabulary}
-\caption{comparison of inner control loop rates among several multirotor MAV FCUs}
-\label{table:mr_loop_rates}
-\end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{RESULTS}
-
-\begin{table}
-\centering
-  \begin{tabulary}{0.5\textwidth}{|C|C|C|C|}
-  \hline
-  message type & message size (bytes) & avg trip time (ms) & max trip time (ms) \\ \hline
-  Commands     & 36                   & 0.117             & 0.2               \\ \hline
-  IMU          & 22                   & 0.115             & 0.19              \\ \hline
-  \end{tabulary}
-  \caption{latency test results for ROSflight}
-  \label{table:latency_results}
-\end{table}
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{fig/hummingbird_att.eps}
-	\includegraphics[width=0.5\textwidth]{fig/hummingbird_vel.eps}
-	\caption{Position, velocity, and attitude estimates and true state plotted as a function of time for a small multirotor MAV.}
-		\label{fig:hummingbird_results}
-\end{figure}
-
-Communication latency between the FCU and onboard computer was tested by sending commands to the FCU where they were registered and then requested from the FCU and returned.
-The send and request cycle was performed at 200~Hz.
-This task required the FCU to record the incoming command, receive a request for its return, and send it, while also operating the control loops at full rate.
-Because response rates may vary slightly due to the order of higher-priority routines in the FCU, 10,000 cycles were performed and the average and maximum cycle times were recorded.
-IMU sensor measurement latency was tested by similarly sending and immediately requesting a message of the same size as the IMU message.
-Results for both test are summarized in table~\ref{table:latency_results}.
-Because I$^2$C communication to the IMU is higher priority than the USB connection to the onboard computer, it can be assumed that actual IMU latency is lower than what was recorded in this test.
-
-Besides IMU, ROSflight supports measurements from a magnetometer, a differential pressure sensor, a barometer and a sonar.
-Because these sensors are used to measure much slower dynamics than accelerometer and gyroscope readings and because total communication bandwidth is limited, these sensors have been integrated and tested at 20~Hz with IMU readings at 150~Hz.
-These speeds are are adjustable if required.
-GPS, cameras, laser scanners and other more complicated sensors used in full state estimation can be connected directly to the onboard computer with a standard USB connections.
-
-Both multirotors shown in figures~\ref{fig:hummingbird} and~\ref{fig:mikey} were flown in a motion capture room under manual control.
-An EKF was implemented on the onboard computer on both MAVs which estimated the position, velocity and attitude of the MAV as well as biases for the accelerometer $x$ and $y$ axes as well as biases for all three gyroscope axes.
-A motion capture system was used to provide position measurements that were fused with IMU for full-state estimation.
-The estimates for velocity and attitude for the smaller quadrotor flight have been plotted alongside their respective motion capture measurements in figure~\ref{fig:hummingbird_results}.
-As can be seen from the results, estimates were able to track effectively throughout the flight.
+constraint sensitivity - aerodynamics, engine efficiencies...
 
@@
 \section{CONCLUSION}
-ROSflight is a new research tool that enables rapid development and testing of autopilot code for MAVs.
-It leverages hardware and software developed by the open-source community to enable accurate attitude estimation and control with a high-bandwidth, low-latency connection to ROS.
-While some of these features exist in current solutions, ROSflight provides state-of-the-art performance for a fraction of the cost of these other solutions while also providing extensive flexibility in platform and onboard computer choice.
+
+lots of really good conclusions
 

### 2016-03-22 python/flight_time.py

@@ -1,0 +1,200 @@
+#!/usr/bin/env python
+import numpy as np
+from math import sqrt, exp, sin, cos, pi
+
+def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_hover, n_rotors, payload, frame, rho_air, E_rhof, n_engine):
+
+#    --- design variables ---
+#    eng_displace    engine displacement (cubic centemeter)
+#    fuel_rate       rate that fuel is consumed (kg/s)
+#    omega         angular velocity of the rotor (rad/s)
+#    R_rotor         rotor radius (meters)
+#    c_rotor         cord of the rotor (meters)
+#    f_cap           fuel capacity (liters)
+#    theta_hover     pitch angel of the blade in hover (rad)
+#    --- parameters ---
+#    n_rotors        number of 2 blade rotors  !!!! this is a discrete variable that can't be dealt with yet but could be later
+#    payload         mass of the payload (kg)
+#    frame           mass of the frame (kg) !!!! this could be calulated from all up weight and radius of the rotor
+#    rho_air         air density (kg/m^3)
+#    E_rhof          energy density of fuel (Joules/kg)
+#    n_engine        efficiency of the engine !!!! this is a param for now but could be calculated from an engine model
+
+    W_motor = eng_displace*0.02145 + 0.25412   #mass of the motor (Kg) !!!! this is an emperical best fit linear relationship
+    P_max = (eng_displace*0.086585514 + 1.357856601)*745.7 #max power (W) as a function of displacment based on emperical best fit linear relationship
+    W_f = f_cap*.75                            #Fuel mass (kg)
+    AUW = W_motor + W_f + frame + payload      #All up mass in kg (about 52 lbs)
+
+#    --- first find a induced air velocity ---  !!! this is an assumption that we will need to improve later
+    T = AUW*9.8/n_rotors                       #Thrust required
+    V_i = sqrt(T/(2*pi*rho_air*R_rotor**2))     #Induced air velocity
+
+#   --- next integrate over the blade to produce T thrust and Q torque ---
[…]

### 2016-03-23 paper.tex
added motivation and figure

@@
 \usepackage{epsfig} % for postscript graphics files
+\usepackage{graphicx}
 \usepackage{mathptmx} % assumes new font selection scheme installed
@@
 
+\makeatletter
+\newenvironment{tablehere}
+  {\def\@captype{table}}
+  {}
+
+\newenvironment{figurehere}
+  {\def\@captype{figure}}
+  {}
+\makeatother
+
 \title{\LARGE \bf
-Awesome quad!!!
+Multi-dimensional Optimization of Gasoline-fueled Variable Pitch Multirotor Aircraft
 }
@@
 
-Hi this is the abstract
+Currently, there are numerous multirotor UAVs that are available commercially and to consumers that are capable of lifting small payloads for an endurance of approximately 30 minutes.
+While these current multirotors are good for short endurance missions, there are few options for larger, higher payload capacity multirotors that are able to maintain flight for longer than an hour.
+This paper describes the optimization process used to develop a multirotor platform that maximizes the flight time of a gasoline-fueled multirotor UAV by varying several design variables related to the power required and the aerodynamics of the model.
 
@@
 
+This paper looks at the why and how we are going to optimize a multirotor platform to maximize flight time
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{itemize}
-	\item{Why a gas quad}
-	\item{figure of current solutions}
-\end{itemize}
+Almost all commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
+Although these batteries are very good and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
+The average specific energy for Lithium-Polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
+This means that a gasoline-fueled multirotor can potential fly much longer than a battery powered multirotor because it can carry more energy onboard.
+
+(Will talk more about some of the details this entails in the paper).
+
+\begin{figurehere}
+	\begin{center}
+		\includegraphics[width=.40\textwidth]{current_capabilities.jpg}
+		\caption{\textit{Graphic showing current platforms and how our would be better.}}
+		\label{current_cap}
+	\end{center}
+\end{figurehere}
 

### 2016-03-23 paper.tex
some small changes
Updated intro

@@
 
+\newcommand{\vect}[1]{\ensuremath{\mathbf{#1}}}
+\newcommand{\mat}[1]{\ensuremath{\mathbf{#1}}}
+\newcommand{\transpose}{\ensuremath{\mathsf{T}}}
+\newcommand{\of}[1]{\ensuremath{\left(#1\right)}}
+
 \title{\LARGE \bf
@@
 
-\author{Daillin Briggs, Gary Ellingson% <-this % stops a space
+\author{Dallin Briggs, Gary Ellingson% <-this % stops a space
 %\thanks{$^{1}$James Jackson is a MS student in the Department of Mechanical Engineering, Brigham Young University
@@
 
-This paper looks at the why and how we are going to optimize a multirotor platform to maximize flight time
+\begin{itemize}
+	\item{This paper looks at the why and how we are going to optimize a multirotor platform to maximize flight time.}
+	\item{Talk about how the variables that are going to be optimized are rotor radius, fuel consumption rate, number of blade per rotor, number of rotors, rotor speed, and rotor pitch.}
+	\item{Talk about why these variables were selected as design variables.}
+	\item{Talk about some physical constraints, especially those imposed by the FAA.}
+\end{itemize}
 
@@
 
-(Will talk more about some of the details this entails in the paper).
+\begin{itemize}
+	\item{Will talk more about some of the details using a gas engine entails in the paper.}
+	\item{Speak briefly on the reasons for requiring variable pitch with a gas engine.} 
+	\item{Speak on current platforms and capabilities and explain how this could fundamentally change the way multirotors are used in commercial applications, but not really for consumers.}
+\end{itemize}
 
@@
 \section{METHODS}
-analysis models
 
-Empirical engine models
-aerodynamic models
-Simplifications - we are stupid
+First model used in the beginning of the project was 
+$$t_{flight} = \frac{E_{cap} \eta_{engine}}{n_{rotors}P_{adj}},$$
+
+\begin{itemize}
+	\item{Analysis models}
+	\begin{itemize}
+		\item{Empirical engine models}
+		\item{Rotor Aerodynamic models}
+	\end{itemize}
+	\item{Simplifications - since we are stupid}
+\end{itemize}
 
 \section{OPTIMIZATION}
-Setup the formal optimization problems
+Setup the formal optimization problem
+\begin{equation}
+min. \quad f\of{x} = -flightTime\of{x}
+\label{eq:objective}
+\end{equation}
+\begin{equation}
+w.r.t. \quad x = [variables]^T
+\label{eq:vars}
+\end{equation}
+\begin{equation}
+s.t. \quad cons\of{x} <= 0 
+\label{eq:constrants}
+\end{equation}
 
 What optimizers, setup, comparing matlab and python.
-figure of Pareto fronts...
+
+Figure of Pareto fronts...
 
 constraint sensitivity - aerodynamics, engine efficiencies...
+
+\section{RESULTS AND DISCUSSION}
+
+\begin{itemize}
+	\item{Table showing the optimal design parameters}
+\end{itemize}
 

### 2016-03-23 paper.tex
added some background and references stuff

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Background}
+
+Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
+There is a variable pitch quad commercially available \cite{stingray2016}.
+There are also several hobbyist attempts at gas powered variable pitch quads \cite{diy2016}, \cite{hackaday2016}.
+
+In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a large payload.
+
+There are many commercially available parts for a gas quad, including motors
 

### 2016-03-24 paper.tex
Updated model

@@
 
-First model used in the beginning of the project was 
-$$t_{flight} = \frac{E_{cap} \eta_{engine}}{n_{rotors}P_{adj}},$$
+Much of the work done in the methodology of optimizing flight time was to find adequate models for the thrust, torque, and power required from the main rotor disks.
+
+First model used in the beginning of the project used a simplified version of Momentum theory and actuator disk theory from \cite{helicopters2016}.
+This model took into account the number of rotors, the rotor radius, the disc loading of each rotor, and how much fuel was available onboard the aircraft.
+However, this model did not take into account the blade speed, blade pitch angle, or the number of blades per rotor.
+The model did, however, give us a baseline for initial optimization results.
+
+Second model used to develop the optimization framework for this project was based on an MIT lecture given on aircraft propeller performance given in \cite{mit2016}.
+In order to test the theory of this model and also to further the progress of the optimization model, only two of the design variables were chosen to optimize: rotor radius and fuel consumption rate, $f_{rate}$.
+Flight time was then calculated by
+\begin{equation}
+	t_{flight} = f_{capacity}/f_{rate}.
+\end{equation}
+This allowed for a very easy calculation in which to test the optimization technique.
+Without constraints, the solution to maximizing flight time would be trivial.
+To properly use the constraints, power produced ($P_{prod}$), power required ($P_{req}$), thrust produced ($T_{prod}$), and thrust required ($T_{req}$) all had to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
+These calculations used some approximations in the rotor dynamics.
+The following calculations come from a MIT lecture on thermodynamics.
+
+In order to calculate thrust of each rotor blade, the thrust has to be calculated differentially along the blade because of the angular velocity. The differential thrust is given by 
+\begin{equation}
+	dT = dLcos(\phi + \alpha_i) - dDsin(\phi + \alpha_i)
+\end{equation}
+
+where $\phi = \tan^{-1}(\frac{V_i}{V_n})$, and $\alpha_i = \theta - \tan^{-1}(\frac{V_i}{V_n})$ and $\theta$ is the blade pitch angle (held constant in this case of the optimization).
+$dL$ and $dD$ are the differentials of the lift and drag, respectively, and are given by
+\begin{equation}
+	dL = \frac{1}{2}\rho V_e^2 c C_l dr
+\end{equation}
+
+and
+
+\begin{equation}
+	dD = \frac{1}{2}\rho V_e^2 c C_d dr.
+\end{equation}
+
+The coefficients of lift and drag were held constant at $C_l = 0.5438$ and $C_d = 0.0141$ (obtained from a symmetric airfoil pitch at $\alpha = 5$ degrees).
+Splitting the rotor into 10 sections and integrating seemed to give a reasonable estimate to the thrust produced by each rotor.
+This was then constrained by making sure the thrust produced was greater than or equal to the thrust required, which was
+\begin{equation}
+	T_{req} = AUW\times g.
+\end{equation}
+
+In order to make sure that enough power was being produced by the engine (and prevent the fuel consumption rate to going to 0), the power produced was given by 
+\begin{equation}
+	P_{prod} = f_{rate}E_{\rho f} \eta_{engine}
+\end{equation}
+
+and was made sure that it was greater than or equal to the power required. To calculate the power required, the torque required to turn each rotor was calculated differentially by the equation
+\begin{equation}
+	dQ = r[dLsin(\phi + \alpha_i) + dDcos(\phi + \alpha_i)].
+\end{equation}
+
+The power required was then calculated by 
+\begin{equation}
+	P_{req} = \omega_rQ_{req}2n_{rotors}.
+\end{equation}
+
+This model did converge to an optimal solution when executed with gradient based optimization methods discussed in the next section.
+However, in this model, not all of the design variables were implemented due to the complexity these additional variables brought to the calculus in this method.
+
+The third model used to try to represent the aerodynamics of each rotor blade was
 

### 2016-03-24 paper.tex
more methods stuff

@@
 
-There are many commercially available parts for a gas quad, including motors
+There are many commercially available parts for a gas quad, including engines and variable pitch helicopter propellers.
 
@@
 \section{METHODS}
+
+Because the quad is going to be build using commercially available off the shell parts the optimization must be closely based on reality.
+Ultimately that means we only can pick the size of engine that we are going to use.
+All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.
+We collected data about popular two stroke RC aircraft engines from several sources to produce the below plots.
+
+\begin{figurehere}
+	\includegraphics[width=0.5\textwidth]{mass.png}
+	\caption{Stand in figure showing the data used to obtain the empirical based surrogate model of engine mass as a function of displacement size.}
+		\label{fig:mass}
+\end{figurehere}
+
+\begin{figurehere}
+	\includegraphics[width=0.5\textwidth]{max_power.png}
+	\caption{Stand in figure showing the data used to obtain the empirical based surrogate model of engine maximum power as a function of displacement size.}
+		\label{fig:power}
+\end{figurehere}
+
+The engine efficiency (friction and heat transfer losses), however, is a function of power capability of the engine (max power) and the power engine is required to produce.
+For this we will need a better model.
+\textit{We are still working on this.
+Since efficiency numbers for RC engines aren't publicly available, we need a scientific model.
+For now we just assume a constant efficiency.
+We plan to work on this in the future.}
 
@@
 
-\begin{itemize}
-	\item{Analysis models}
-	\begin{itemize}
-		\item{Empirical engine models}
-		\item{Rotor Aerodynamic models}
-	\end{itemize}
-	\item{Simplifications - since we are stupid}
-\end{itemize}
+\section{OPTIMIZATION}
 
-\section{OPTIMIZATION}
 Setup the formal optimization problem

### 2016-03-24 paper.tex
added nasa model

@@
 
-The third model used to try to represent the aerodynamics of each rotor blade was
+The third model used to represent the aerodynamics and thrust of each rotor blade was from \cite{chen1979simplified}, a NASA model used in helicopter simulations several years ago.
+Instead of trying to implement blade element theory across the blade, this new model is able to approximate the thrust and torque required for the rotor given numerous variables.
+Instead of only varying rotor radius and fuel consumption rate, like last homework's model, this model attempts to optimize given rotor radius, fuel consumption rate, number of blade per rotor, number of rotors, rotor speed, and rotor pitch.
+The thrust, T, is given by equation \ref{thrust} and the torque, Q, is represented by equations \ref{torque} and \ref{torquecoeff}.
+
+\begin{multline}
+T = \frac{N}{2} \rho a c R (\omega R)^2 (.5 (1-\epsilon^2) \lambda \\
+ + \theta_0 (\frac{1}{3} + \frac{\mu^2}{2} (1-\epsilon)) + \theta_t (.25+\frac{\mu^2}{2} (1-\epsilon^2))\\
+- \frac{\mu}{2} (1-\epsilon^2) (B_{lc}-K_l b_l) - a_0 (\frac{1}{3} +\frac{\mu^2}{2} (1-\epsilon)) K_l \\
++ a_l (\frac{\mu}{2} \epsilon (1-\epsilon)) - \frac{\dot{a_0}}{\omega} (\frac{1}{3} -\frac{\epsilon}{2}) + \frac{\dot{b_l}}{\omega} (\frac{\mu}{4} (1-\epsilon)^2) \\
++ \frac{\mu}{4} (1-\epsilon^2) (\frac{p}{\omega cos(\beta_w)} + \frac{q}{\omega sin(\beta_w))} ) - N M_\beta \ddot{a_0}
+\label{thrust}
+\end{multline}
+and
+\begin{equation}
+Q = N/2 \rho a c R^2 (\omega R)^2 (Q_{coeff})
+\label{torque}
+\end{equation}
+where
+\begin{multline}
+Q_{coeff} = \delta/(4 a) (1 + (1-\epsilon^2) \mu^2) \\
+ - (\theta_0 - K_l a_0)(\lambda/3+(\epsilon/3-.25) \dot{a_0}/\omega + \mu/6 \dot{b_l}/\omega\\  
+- \mu \epsilon/4 (\dot{b_l}/\omega-a_l) + \mu/6 (p/\omega cos(\beta_w) +q/\omega sin(\beta_w))) \\
+- (A_{lc}-K_l a_l) ((.125-\epsilon/6) (\dot{a_l}/\omega+b_l)-\mu/6 a_0 \\
++b_l/16 (1-\epsilon^2) \mu^2+1/8 (-p/\omega sin(\beta_w)+q/\omega cos(\beta_w)))\\ 
+- (B_lc-K_l b_l) ((.125-\epsilon/6) (\dot{b_l}/\omega-a_l)\\
++(\epsilon/4-1/6) \mu \dot{a_0}/\omega+.5 (1-\epsilon^2) (\mu \lambda/2+a_l/8 \mu^2)\\ 
++ 1/8 (p/\omega cos(\beta_w)+q/\omega sin(\beta_w))) - \theta_t (\lambda/4\\ 
++ (\epsilon/4-1/5) \dot{a_0}/\omega+\mu/8 (\dot{b_l}/\omega)-\epsilon \mu/6 (\dot{b_l}/\omega-a_l))\\  
+- .5 (1-\epsilon^2) (\lambda^2+\lambda \mu a_l+2 \lambda \epsilon \dot{a_0}/\omega\\  
++\mu \epsilon (a_l \dot{a_0}/\omega+a_0 (\dot{a_l}/\omega+b_l))\\  
++ \mu^2 (a_0^2/2+3/8 a_l^2+1/8 b_l^2)) + \mu/3 (a_l (\dot{a_0}/\omega) \\
++ a_0 (\dot{a_l}/\omega+b_l))+2/3 \lambda (\dot{a_l}/\omega)\\  
+- (-\mu/3 a_0 + (.25-\epsilon/3) (\dot{a_l}/\omega+b_l)) \\ 
+(-p/\omega sin(\beta_w)+q/\omega cos(\beta_w))\\ 
+- (.25-\epsilon/3) (\dot{b_l}/\omega-a_l) (p/\omega cos(\beta_w)\\ 
++ q/\omega sin(\beta_w)) - .125 (-p/\omega sin(\beta_w)+q/\omega cos(\beta_w))^2 \\ 
+-.125 (p/\omega cos(\beta_w)+q/\omega sin(\beta_w))^2 -  \\
+(.25-2/3 \epsilon+\epsilon^2/2) ((\dot{a_0}/\omega)^2 \\ 
++ .5 ((\dot{a_l}/\omega+b_l)^2+(\dot{b_l}/\omega-a_l)^2))
+\label{torquecoeff}
+\end{multline}
+
+Please see \cite{chen1979simplified} for definition of terms.
+Although this model takes into account all of the design variables chosen for this project, it is difficult to optimize because of the complexity of the functions.
+\textit{Currently, this model fails to converge to a solution, so we are working with it to find the source of the issue.}
 

### 2016-03-24 paper.tex
optimization section work

@@
 
-Because the quad is going to be build using commercially available off the shell parts the optimization must be closely based on reality.
-Ultimately that means we only can pick the size of engine that we are going to use.
+Because the quad is going to be build using commercially available off-the-shelf parts the optimization must be closely based on reality.
+This means we only can pick the size of engine that we are going to use.
 All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.
@@
 
+Although the off-the-shelf constraint also applies to choosing the rotor parameters, continuous variables approximating the rotor specs are used and then analyzed as follows.
+
 Much of the work done in the methodology of optimizing flight time was to find adequate models for the thrust, torque, and power required from the main rotor disks.
@@
 In order to test the theory of this model and also to further the progress of the optimization model, only two of the design variables were chosen to optimize: rotor radius and fuel consumption rate, $f_{rate}$.
-Flight time was then calculated by
-\begin{equation}
-	t_{flight} = f_{capacity}/f_{rate}.
-\end{equation}
-This allowed for a very easy calculation in which to test the optimization technique.
-Without constraints, the solution to maximizing flight time would be trivial.
-To properly use the constraints, power produced ($P_{prod}$), power required ($P_{req}$), thrust produced ($T_{prod}$), and thrust required ($T_{req}$) all had to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
+
+To properly analyses the system, power required ($P_{req}$) and thrust produced ($T_{prod}$) have to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
 These calculations used some approximations in the rotor dynamics.
@@
 
-What optimizers, setup, comparing matlab and python.
+Flight time was calculated by 
+\begin{equation}
+	t_{flight} = K_{energy}*f_{capacity}/f_{rate}.
+\end{equation}
+Without constraints, the solution to maximizing flight time would be trivial and impossible.
+To optimize the system we use the constraints on the power produced ($P_{prod}$), power required ($P_{req}$), thrust produced ($T_{prod}$), and lift required ($T_{req}$) all had to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
 
-Figure of Pareto fronts...
+We then will discuss the following:
+\begin{itemize}
+	\item{scaling of the design variables}
+	\item{how we got gradients}
+	\item{the optimization method that we utilized}
+\end{itemize}
 
-constraint sensitivity - aerodynamics, engine efficiencies...
+So far our problem has been implemented both in python and in matlab.  Here we can talk about the benefits and challenges of using each.
+
+Since we can only optimize for maximum flight time give a payload (parameter) we can further explore how the characteristics of a gas quad change as we optimize the flight time for different payloads.
+It may turn out that the optimal quad parameters change greatly as it is optimized for different flight times.
+In this case we could choose several payloads and optimize over all of then together.
+\textit{We will present these results with a figure and some discussion.}
+
+\begin{figurehere}
+	\includegraphics[width=0.5\textwidth]{payload.jpg}
+	\caption{Stand in figure showing optimal flight time as a function of payload.}
+		\label{fig:payload}
+\end{figurehere}
+
+Here we can do some discussion on constraint sensitivity - aerodynamics, engine efficiencies...
 
@@
 
-\begin{itemize}
-	\item{Table showing the optimal design parameters}
-\end{itemize}
+We have found that the flight time of a gas-quad can be maximized by constructing with the following design.
+
+\begin{tablehere}
+\centering
+\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
+\hline
+       & displacement & fuel rate & rotor radius & other \\ \hline
+  & 100 & 30 & 250 & 150 \\ \hline
+\end{tabulary}
+\caption{Optimal design variables}
+\label{table:fw_loop_rates}
+\end{tablehere}
 

### 2016-03-24 paper.tex
design variables

@@
 
+Our design variables are rotor radius, fuel consumption rate, number of blades per rotor, number of rotors, rotor speed, and rotor pitch.
+Rotor radius and number of blades per rotor, and number of rotors were chosen because this is an easy parameter to vary on multirotors due to the availability of numerous rotor blades available for small rotor aircraft.
+These variables also have a large influence on the flight time of the multirotor.
+Rotor speed and rotor pitch were chosen because these variables are easy to change on the physical system through selecting the proper gearing ratios.
+Fuel consumption rate seems like an odd variable for an optimizer because intuitively an optimizer would just minimize this value to maximize the flight time; however, fuel consumption rate is constrained by the size of engine required to produce the amount of power needed to provide enough lift for the multirotor.
+All of these design variables, however, have bound constraints that keep them in a reasonable space of commercially available parts/performance.
+
 Setup the formal optimization problem

### 2016-03-24 paper.tex
final outline assignment stuff

@@
 Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
-There is a variable pitch quad commercially available \cite{stingray2016}.
+There is a small, electric variable pitch quad commercially available \cite{stingray2016}.
 There are also several hobbyist attempts at gas powered variable pitch quads \cite{diy2016}, \cite{hackaday2016}.
 
-In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a large payload.
+In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload.
 
@@
 
-Almost all commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
+All commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
 Although these batteries are very good and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
@@
 \begin{equation}
-min. \quad f\of{x} = -flightTime\of{x}
+min. \quad f\of{x} = -t_{flight}\of{x}
 \label{eq:objective}
@@
 \begin{equation}
-s.t. \quad cons\of{x} <= 0 
+s.t. \quad cons\of{x} \leq 0 
 \label{eq:constrants}
@@
 \end{equation}
-Without constraints, the solution to maximizing flight time would be trivial and impossible.
-To optimize the system we use the constraints on the power produced ($P_{prod}$), power required ($P_{req}$), thrust produced ($T_{prod}$), and lift required ($T_{req}$) all had to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
+Without constraints, the solution to maximizing flight time would be trivial to calculate as well as impossible to construct.
+Therefore to optimize the system we must use constraints that ensure the quad is feasible.
+The constrains are on the power produced by the engine ($P_{prod}$), power required by the prop ($P_{req}$), thrust produced by the prop ($T_{prod}$), lift required by the quad ($T_{req}$) and maximum power capabilities of the engine ($P_{max}$).
+Where $P_{preq} \leq P_{prod}$, $T_{req} \leq T_{prod}$, and $P_{prod} \leq P_{max}$.
 
@@
 	\item{the optimization method that we utilized}
+	\item{python vs matlab benefits and challenges}
 \end{itemize}
 
-So far our problem has been implemented both in python and in matlab.  Here we can talk about the benefits and challenges of using each.
-
 Since we can only optimize for maximum flight time give a payload (parameter) we can further explore how the characteristics of a gas quad change as we optimize the flight time for different payloads.
-It may turn out that the optimal quad parameters change greatly as it is optimized for different flight times.
+It may turn out that the optimal quad parameters change greatly as it is optimized for different payload.
 In this case we could choose several payloads and optimize over all of then together.
@@
 
-Here we can do some discussion on constraint sensitivity - aerodynamics, engine efficiencies...
+%Here we can do some discussion on constraint sensitivity - aerodynamics, engine efficiencies...
 

### 2016-04-06 thrust.m
Added thrust model

@@ -1,0 +1,78 @@
+function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
+
+    % This method of calculating thrust uses blade element theory
+    %x consists of the number of blades in each rotor, rotor radius, 
+    %angular velocity, number of rotors, and blade pitch, fuel consumption rate.
+%     keyboard
+    N = x(1);
+    R = x(2);
+    omega = x(3);
+    n = x(4);
+    theta = x(5);
+    f_rate = x(6)/10^3;
+
+    V_c = 0;                % Vertical velocity is zero because of hover.
+    a = .1091;               % Lift curve slope for naca 0012
+    rho = 1.225;            % Air density, kg/m^3
+    C = .09*R;              % Chord length, based on rc rotor blades (extrapolated).
+    T = 0;
+    Q = 0;
+    P = 0;
+    n_engine = 0.15;                            % Efficiency of the engine
+    E_rhof = 32.4e6;                            %Energy density of fuel (Joules/L)
+    f_cap = 19;             % In liters.
+    W_motor = 2.1;                              %Motor weight
+    W_f = f_cap*.75;                            %Fuel weight (kg)
+    payload = 5;                                %Payload weight
+    W_frame = 7;                                %Frame weight
+    AUW = W_motor + W_f + W_frame + payload;    %All up weight in kg (about 52 lbs)
+    
+    m = 10;
+    for i=1:m
+        dr = R/m;
+        r = i*R/m-.5*R/m;
+        s = n*C/pi*r;
+        phi = 1/16*(-sqrt(a)*sqrt(s)*sqrt(a*s+32*theta)-a*s); %atan(U_P/U_T);          % Small angle approximation because U_P << U_T.
+        alpha = theta - phi;
+        C_l = -98.031*alpha^3 + 0.9481*alpha^2 + 8.2078*alpha - 0.009;          % Lift coefficient based on Xfoil data.
+        C_d = 51.031*alpha^4 - 0.9405*alpha^3 - 0.5316*alpha^2 + 0.0191*alpha + 0.0168;   % Drag coefficient based on xfoil data.
+        t_c = C_l/6;
+        lambda_i = sqrt(s*t_c/2);
+        v_i = omega*r*lambda_i;
+        U_T = omega*r;          % Tangential component of velocity.
[…]

### 2016-04-06 python/project.py
python optimizer added, matlab file moved

@@ -1,0 +1,48 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from math import sqrt, exp, sin, cos
+import time
+from scipy.optimize import minimize
+from flight_time import obj_func
+
+
+if __name__ == '__main__':
+
+    from pyoptwrapper import optimize
+    from pyoptsparse import NSGA2, SNOPT, NLPQLP, ALPSO, SLSQP
+
+    eng_displace = 80.0
+    fuel_rate = .02
+    omega_r = 1500*2*3.14159/60 #rpm to rad/s
+    R_rotor = 0.705
+    c_rotor = 0.05
+    fuel_cap = 1.0
+    theta_hover = 20*3.14159/180
+    #    x0 = [eng_displace, fuel_rate, omega_r, R_rotor, c_rotor, fuel_cap, theta_hover]
+    x0 = [1., 1., 1., 1., 1., 1., 1.]
+    #    x0 = [fuel_rate, R_rotor, theta_hover]
+
+    lb = [0, 0, 0, 0, 0, 0, 0]
+    ub = [1200.0/80.0, 1.0/.02, 7500.0/1500.0, 2.0/0.705, 0.1/0.05, 12.0/1.0, 30.0/20.0]
+
+#    x0 = [3.55613094,   0.24996519,   0.72821006,   1.27615376,   1.53249666,  12.,
+#    1.32161584]
+#    x0 = [  9.85947913e-04,   1.73583194e-01,   4.12512324e-01 ,  1.84357306e+00,
+#    1.75334613e+00,   1.19999999e+01,   6.94509532e-01]
+    x0 = [  3.37894914e-05,   2.46734305e-01,   3.92570916e-01,   1.76376878e+00,
+    1.24587979e+00,   1.19999995e+01,   1.47021053e+00]
+
+    print obj_func(x0)
+
+#    optimizer = SNOPT()
+    optimizer = NSGA2()
+    optimizer.setOption('maxGen', 200)
+
+    xopt, fopt, info = optimize(obj_func, x0, lb, ub, optimizer)
+    print ub
+    print 'SNOPT:', xopt, fopt, info
+
+    print obj_func(xopt)
+
+
+

### 2016-04-06 python/engine_models.py
gradient based aproaches are working, improved engine models

@@ -1,0 +1,69 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+import matplotlib.pyplot as plt
+import numpy as np
+
+            #name :  [displacment(cc), mass(kg), power(hp)]
+Engines = {'DA-35' : [35, 0.935, -1],
+          'DA-50' : [50, 1.33, -1],
+          'DA-60' : [60.5, 1.41, -1],
+          'DA-70' : [70, 1.61, -1],
+          'DA-85' : [85.9, 1.95, -1],
+          'DA-100L' : [100, 2.53, 9.8],
+          'DA-100inline' : [100, 3.12, -1],
+          'DA-120' : [121, 2.25, -1],
+          'DA-150L' : [150, 3.3, -1],
+          'DA-170' : [171.5, 3.56, -1],
+          'DA-200' : [200, 4.95, 19],
+          'Evilution NX' : [8.52, 0.435, -1],
+          'OS 50SX-H' : [8.17, 0.406, 1.87],
+          'OS 105HZ-R' : [17.17, 0.608, 3.75],
+          'HK MLD-70' : [70, 2.18, 6.6]}
+
+
+data = np.array(Engines.values())
+f = data[:,1]
+A = np.array([data[:,0], np.ones(f.size)])
+
+a_w, b_w  = np.linalg.lstsq(A.T,f)[0]
+#print a,b
+plt.figure()
+plt.plot(data[:,0], data[:,1], 'ko')
+plt.plot([0,200], [b_w,a_w*200+b_w], '-k')
+#plt.show()
+
+powerdata = np.array([0])
+for row in data[:]:
+    if row[2] > 0 and powerdata.size < 3:
+        powerdata = np.array(row)
+    elif row[2] > 0:
+        powerdata = np.append(powerdata, row)
+
+powerdata = powerdata.reshape(5,3)
+f = powerdata[:,2]*745.699872 #put the powerdata in Watts
+
+plt.figure(2)
+# data points
+plt.plot(powerdata[:,0], f, 'ko')
+
+# linear model
+A = np.array([powerdata[:,0], np.ones(f.size)])
+a, b  = np.linalg.lstsq(A.T,f)[0]
+#print a,b
+plt.plot([0,200], [b,a*200+b], '-k')
+
+# data from http://www.barnardmicrosystems.com/UAV/engines/2_stroke.html
+a_ref = 76.0
[…]

### 2016-04-06 python/flight_time.py

@@ -2,12 +2,14 @@
 import numpy as np
 from math import sqrt, exp, sin, cos, pi
 
+prt = False
+
 def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_hover, n_rotors, payload, frame, rho_air, E_rhof, n_engine):
 
 #    --- design variables ---
 #    eng_displace    engine displacement (cubic centemeter)
 #    fuel_rate       rate that fuel is consumed (kg/s)
-#    omega         angular velocity of the rotor (rad/s)
+#    omega           angular velocity of the rotor (rad/s)
 #    R_rotor         rotor radius (meters)
 #    c_rotor         cord of the rotor (meters)
 #    f_cap           fuel capacity (liters)
@@ -21,7 +23,8 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
 #    n_engine        efficiency of the engine !!!! this is a param for now but could be calculated from an engine model
 
     W_motor = eng_displace*0.02145 + 0.25412   #mass of the motor (Kg) !!!! this is an emperical best fit linear relationship
-    P_max = (eng_displace*0.086585514 + 1.357856601)*745.7 #max power (W) as a function of displacment based on emperical best fit linear relationship
+    P_max = eng_displace*76.0 + 502.0 #max power (W) as a function of displacment based on emperical best fit linear relationship
+    P_max = -0.0454790393518*eng_displace**2 + 79.5042893755*eng_displace #max power (W) as a function of displacment, best fit linear quadradic
     W_f = f_cap*.75                            #Fuel mass (kg)
     AUW = W_motor + W_f + frame + payload      #All up mass in kg (about 52 lbs)
 
@@ -71,7 +74,14 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
 
     PowerRequired = 8*Q*omega
     PowerProduced = fuel_rate*E_rhof*n_engine
-    c = np.array([PowerProduced - PowerRequired, T - AUW*9.8 , PowerProduced - P_max])
[…]

### 2016-04-06 python/project.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 from math import sqrt, exp, sin, cos
 import time
 from scipy.optimize import minimize
-from flight_time import obj_func
+from flight_time import obj_func, obj_func_print, scale, unscale
 
 
 if __name__ == '__main__':
@@ -11,38 +11,28 @@ if __name__ == '__main__':
     from pyoptwrapper import optimize
     from pyoptsparse import NSGA2, SNOPT, NLPQLP, ALPSO, SLSQP
 
-    eng_displace = 80.0
-    fuel_rate = .02
-    omega_r = 1500*2*3.14159/60 #rpm to rad/s
-    R_rotor = 0.705
-    c_rotor = 0.05
-    fuel_cap = 1.0
-    theta_hover = 20*3.14159/180
-    #    x0 = [eng_displace, fuel_rate, omega_r, R_rotor, c_rotor, fuel_cap, theta_hover]
-    x0 = [1., 1., 1., 1., 1., 1., 1.]
-    #    x0 = [fuel_rate, R_rotor, theta_hover]
+    eng_displace = 80.0 # cc
+    fuel_rate = .02 # kg/s
+    omega_r = 1500 # rpm
+    R_rotor = 0.705 # m
+    c_rotor = 0.05 # m
+    fuel_cap = 1.0 # kg
+    theta_hover = 20 # deg
+    x0 = scale([eng_displace, fuel_rate, omega_r, R_rotor, c_rotor, fuel_cap, theta_hover])
 
     lb = [0, 0, 0, 0, 0, 0, 0]
-    ub = [1200.0/80.0, 1.0/.02, 7500.0/1500.0, 2.0/0.705, 0.1/0.05, 12.0/1.0, 30.0/20.0]
+    ub = scale([1200.0, 1.0, 7500.0, 1.0, 0.1, 12.0, 30.0])
 
-#    x0 = [3.55613094,   0.24996519,   0.72821006,   1.27615376,   1.53249666,  12.,
-#    1.32161584]
-#    x0 = [  9.85947913e-04,   1.73583194e-01,   4.12512324e-01 ,  1.84357306e+00,
-#    1.75334613e+00,   1.19999999e+01,   6.94509532e-01]
-    x0 = [  3.37894914e-05,   2.46734305e-01,   3.92570916e-01,   1.76376878e+00,
-    1.24587979e+00,   1.19999995e+01,   1.47021053e+00]
+#    print obj_func_print(x0)
 
[…]

### 2016-04-07 python/flight_time.py
added a frame model and worked on cleaning things up by scaling apropriatly

@@ -18,6 +18,7 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
 #    n_rotors        number of 2 blade rotors  !!!! this is a discrete variable that can't be dealt with yet but could be later
 #    payload         mass of the payload (kg)
 #    frame           mass of the frame (kg) !!!! this could be calulated from all up weight and radius of the rotor
+    frame = 4.22659612552*R_rotor
 #    rho_air         air density (kg/m^3)
 #    E_rhof          energy density of fuel (Joules/kg)
 #    n_engine        efficiency of the engine !!!! this is a param for now but could be calculated from an engine model
@@ -53,8 +54,10 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
 
         sigma_a = (1 + exp(-(M*(alpha - alpha0))) + exp((M*(alpha + alpha0))))/((1 + exp(-(M*(alpha - alpha0))))*(1 + exp((M*(alpha + alpha0))))) # to model stall
         CL_a = (1 - sigma_a)*(CL_0 + CL_alpha*alpha) + sigma_a*(2*np.sign(alpha)*pow(sin(alpha),2.0)*cos(alpha))
+#        CL_a = (1 - sigma_a)*(-98.031*alpha**3 + 0.9481*alpha**2 + 8.2078*alpha - 0.009) + sigma_a*(2*np.sign(alpha)*pow(sin(alpha),2.0)*cos(alpha))
         AR = 2*R_rotor/c_rotor
         CD_a = CD_p + ((pow((CL_0 + CL_alpha*(alpha)),2.0))/(3.14159*0.9*AR))
+#        CD_a = 51.031*alpha**4 - 0.9405*alpha**3 - 0.5316*alpha**2 + 0.0191*alpha + 0.0168
 
 #        differential drag and lift
         dL = 0.5*rho_air*(Ve**2)*c_rotor*CL_a*dr
@@ -95,7 +98,7 @@ def obj_func(x):
     theta_hover = theta_hover*3.14159/180 # to rad
 
     n_rotors = 6
-    payload = 3.0
+    payload = 6.0
     frame = 2.0
     rho_air = 1.225
     E_rhof = 44.4e6

### 2016-04-07 python/project.py

@@ -21,7 +21,7 @@ if __name__ == '__main__':
     x0 = scale([eng_displace, fuel_rate, omega_r, R_rotor, c_rotor, fuel_cap, theta_hover])
 
     lb = [0, 0, 0, 0, 0, 0, 0]
-    ub = scale([1200.0, 1.0, 7500.0, 1.0, 0.1, 12.0, 30.0])
+    ub = scale([1200.0, 1.0, 7500.0, 3.0, 0.1, 12.0, 30.0])
 
 #    print obj_func_print(x0)
 

### 2016-04-08 thrust.m
Fixed thrust calc

@@ -5,10 +5,10 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
     %angular velocity, number of rotors, and blade pitch, fuel consumption rate.
 %     keyboard
     N = x(1);
-    R = x(2);
-    omega = x(3);
+    R = x(2)/10;
+    omega = x(3)*100;
     n = x(4);
-    theta = x(5);
+    theta = x(5)/10;
     f_rate = x(6)/10^3;
 
     V_c = 0;                % Vertical velocity is zero because of hover.
@@ -47,6 +47,7 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
 %         dT = N*L*dr;
         dT = .5*rho*a*N*omega^2.*C.*(theta-phi).*r.^2.*dr;
 %         dQ = N*(L*phi+D)*r*dr;
+%         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr;
         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr;
         dP = dQ*omega;
         T = dT*n + T;

### 2016-04-08 thrust.m
Fixed thrust calculation

@@ -1,0 +1,79 @@
+function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
+
+    % This method of calculating thrust uses blade element theory
+    %x consists of the number of blades in each rotor, rotor radius, 
+    %angular velocity, number of rotors, and blade pitch, fuel consumption rate.
+%     keyboard
+    N = x(1);
+    R = x(2)/10;
+    omega = x(3)*100;
+    n = x(4);
+    theta = x(5)/10;
+    f_rate = x(6)/10^3;
+
+    V_c = 0;                % Vertical velocity is zero because of hover.
+    a = .1091;               % Lift curve slope for naca 0012
+    rho = 1.225;            % Air density, kg/m^3
+    C = .09*R;              % Chord length, based on rc rotor blades (extrapolated).
+    T = 0;
+    Q = 0;
+    P = 0;
+    n_engine = 0.15;                            % Efficiency of the engine
+    E_rhof = 32.4e6;                            %Energy density of fuel (Joules/L)
+    f_cap = 19;             % In liters.
+    W_motor = 2.1;                              %Motor weight
+    W_f = f_cap*.75;                            %Fuel weight (kg)
+    payload = 5;                                %Payload weight
+    W_frame = 7;                                %Frame weight
+    AUW = W_motor + W_f + W_frame + payload;    %All up weight in kg (about 52 lbs)
+    
+    m = 10;
+    for i=1:m
+        dr = R/m;
+        r = i*R/m-.5*R/m;
+        s = n*C/pi*r;
+        phi = 1/16*(-sqrt(a)*sqrt(s)*sqrt(a*s+32*theta)-a*s); %atan(U_P/U_T);          % Small angle approximation because U_P << U_T.
+        alpha = theta - phi;
+        C_l = -98.031*alpha^3 + 0.9481*alpha^2 + 8.2078*alpha - 0.009;          % Lift coefficient based on Xfoil data.
+        C_d = 51.031*alpha^4 - 0.9405*alpha^3 - 0.5316*alpha^2 + 0.0191*alpha + 0.0168;   % Drag coefficient based on xfoil data.
+        t_c = C_l/6;
+        lambda_i = sqrt(s*t_c/2);
+        v_i = omega*r*lambda_i;
[…]

### 2016-04-08 presentation/presentation.tex
added presentation stuff

\documentclass{beamer}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\mode<presentation>
\usetheme[footline=empty]{Magicc}

\usepackage{graphicx}
\graphicspath{{figures/}}
\DeclareGraphicsExtensions{.pdf, .png, .jpg}

\usepackage{tikz}

% add logo to slide header
\addtobeamertemplate{frametitle}{}{%
\begin{tikzpicture}[remember picture,overlay]
\node[anchor=north east,yshift=2pt] at (current page.north east) {\includegraphics[height=0.75cm]{logo_white}};
\end{tikzpicture}}

% create outline slide at each \section{} (optional)
\AtBeginSection[]{\frame{\frametitle{Outline}\tableofcontents[currentsection]}}

% presentation info
\title[Short Title]{GAS Quad}
\author{Smart Person}
\institute[MAGICC Lab]{\includegraphics[height=0.5in]{logo_gray}}
\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Presentation Guidelines}

You will sign up for a 10 minute presentation time slot. The presentation is graded only for completion. Group projects will have 8 minutes to present and 2 minutes for Q\&A. Individuals will have 6 minutes to present and 2 minutes for Q\&A. Obviously you will not have a lot of time to spend on details. Instead, you should focus on motivation (what the problem is and why it is important), the solution approach, and the most important results. Each member of the team should have a role in the presentation.

\end{frame}

\end{document}

### 2016-04-08 thrust.m

@@ -1,79 +1,0 @@
-function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
-
-    % This method of calculating thrust uses blade element theory
-    %x consists of the number of blades in each rotor, rotor radius, 
-    %angular velocity, number of rotors, and blade pitch, fuel consumption rate.
-%     keyboard
-    N = x(1);
-    R = x(2)/10;
-    omega = x(3)*100;
-    n = x(4);
-    theta = x(5)/10;
-    f_rate = x(6)/10^3;
-
-    V_c = 0;                % Vertical velocity is zero because of hover.
-    a = .1091;               % Lift curve slope for naca 0012
-    rho = 1.225;            % Air density, kg/m^3
-    C = .09*R;              % Chord length, based on rc rotor blades (extrapolated).
-    T = 0;
-    Q = 0;
-    P = 0;
-    n_engine = 0.15;                            % Efficiency of the engine
-    E_rhof = 32.4e6;                            %Energy density of fuel (Joules/L)
-    f_cap = 19;             % In liters.
-    W_motor = 2.1;                              %Motor weight
-    W_f = f_cap*.75;                            %Fuel weight (kg)
-    payload = 5;                                %Payload weight
-    W_frame = 7;                                %Frame weight
-    AUW = W_motor + W_f + W_frame + payload;    %All up weight in kg (about 52 lbs)
-    
-    m = 10;
-    for i=1:m
-        dr = R/m;
-        r = i*R/m-.5*R/m;
-        s = n*C/pi*r;
-        phi = 1/16*(-sqrt(a)*sqrt(s)*sqrt(a*s+32*theta)-a*s); %atan(U_P/U_T);          % Small angle approximation because U_P << U_T.
-        alpha = theta - phi;
-        C_l = -98.031*alpha^3 + 0.9481*alpha^2 + 8.2078*alpha - 0.009;          % Lift coefficient based on Xfoil data.
-        C_d = 51.031*alpha^4 - 0.9405*alpha^3 - 0.5316*alpha^2 + 0.0191*alpha + 0.0168;   % Drag coefficient based on xfoil data.
-        t_c = C_l/6;
-        lambda_i = sqrt(s*t_c/2);
-        v_i = omega*r*lambda_i;
[…]

### 2016-04-08 python/flight_time.py
removed blade chord as design variable and started adding an other aerodynamic model

@@ -4,7 +4,7 @@ from math import sqrt, exp, sin, cos, pi
 
 prt = False
 
-def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_hover, n_rotors, payload, frame, rho_air, E_rhof, n_engine):
+def flight_time(eng_displace, fuel_rate, omega, R_rotor, f_cap, theta_hover, n_rotors, payload, frame, rho_air, E_rhof, n_engine):
 
 #    --- design variables ---
 #    eng_displace    engine displacement (cubic centemeter)
@@ -12,6 +12,7 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
 #    omega           angular velocity of the rotor (rad/s)
 #    R_rotor         rotor radius (meters)
 #    c_rotor         cord of the rotor (meters)
+    c_rotor = 0.09*R_rotor
 #    f_cap           fuel capacity (liters)
 #    theta_hover     pitch angel of the blade in hover (rad)
 #    --- parameters ---
@@ -28,6 +29,88 @@ def flight_time(eng_displace, fuel_rate, omega, R_rotor, c_rotor, f_cap, theta_h
     P_max = -0.0454790393518*eng_displace**2 + 79.5042893755*eng_displace #max power (W) as a function of displacment, best fit linear quadradic
     W_f = f_cap*.75                            #Fuel mass (kg)
     AUW = W_motor + W_f + frame + payload      #All up mass in kg (about 52 lbs)
+
+
+    T, Q = simple_aero_model(AUW, omega, R_rotor, c_rotor, theta_hover, rho_air, n_rotors)
+#    T, Q = other_aero_model(omega, R_rotor, c_rotor, theta_hover, rho_air, n_rotors)
+#    T, Q = nasa(2.0, R_rotor, omega, n_rotors, theta_hover, AUW, rho_air)
+
+    PowerRequired = 8*Q*omega
+    PowerProduced = fuel_rate*E_rhof*n_engine
+    c = np.array([PowerProduced - PowerRequired, T - AUW*9.8 , P_max - PowerProduced, 25 - AUW])
+    global prt
+    if prt == True:
+        print 'PP', PowerProduced
[…]

### 2016-04-08 python/project.py

@@ -15,13 +15,14 @@ if __name__ == '__main__':
     fuel_rate = .02 # kg/s
     omega_r = 1500 # rpm
     R_rotor = 0.705 # m
-    c_rotor = 0.05 # m
     fuel_cap = 1.0 # kg
     theta_hover = 20 # deg
-    x0 = scale([eng_displace, fuel_rate, omega_r, R_rotor, c_rotor, fuel_cap, theta_hover])
+    x0 = scale([eng_displace, fuel_rate, omega_r, R_rotor, fuel_cap, theta_hover])
+    x0 = [ 4.3441622,   0.16620041,  0.10885692,  2.78460696,  6.76453586, 0.85782237]
+    x0 = [ 4.61280557,  0.17376677,  0.08200489,  3.4691025,   7.32432873, 0.86800068]
 
-    lb = [0, 0, 0, 0, 0, 0, 0]
-    ub = scale([1200.0, 1.0, 7500.0, 3.0, 0.1, 12.0, 30.0])
+    lb = [0, 0, 0, 0, 0, 0]
+    ub = scale([1200.0, 1.0, 7500.0, 3.0, 12.0, 30.0])
 
 #    print obj_func_print(x0)
 
@@ -30,9 +31,6 @@ if __name__ == '__main__':
 #    optimizer.setOption('maxGen', 200)
 
     xopt, fopt, info = optimize(obj_func, x0, lb, ub, optimizer)
-    print 'SNOPT:', fopt, info
+    print 'SNOPT:', xopt, fopt, info
 
     out = obj_func_print(xopt)
-
-
-

### 2016-04-08 python/engine_models.py
updated some figures

@@ -28,8 +28,12 @@ A = np.array([data[:,0], np.ones(f.size)])
 a_w, b_w  = np.linalg.lstsq(A.T,f)[0]
 #print a,b
 plt.figure()
-plt.plot(data[:,0], data[:,1], 'ko')
-plt.plot([0,200], [b_w,a_w*200+b_w], '-k')
+plt.xlabel('Engine Size (cc)')  # labels for axes
+plt.ylabel('Engine Mass (kg)')
+plt.title('Mass Model')
+plt.plot(data[:,0], data[:,1], 'ko', label='Data')
+plt.plot([0,200], [b_w,a_w*200+b_w], '-k', label='Best Fit')
+plt.legend(loc='upper left')
 #plt.show()
 
 powerdata = np.array([0])
@@ -43,19 +47,22 @@ powerdata = powerdata.reshape(5,3)
 f = powerdata[:,2]*745.699872 #put the powerdata in Watts
 
 plt.figure(2)
+plt.xlabel('Engine Size (cc)')  # labels for axes
+plt.ylabel('Engine Power (W)')
+plt.title('Power Model')
 # data points
-plt.plot(powerdata[:,0], f, 'ko')
+plt.plot(powerdata[:,0], f, 'ko', label='Data')
 
 # linear model
 A = np.array([powerdata[:,0], np.ones(f.size)])
 a, b  = np.linalg.lstsq(A.T,f)[0]
 #print a,b
-plt.plot([0,200], [b,a*200+b], '-k')
+plt.plot([0,200], [b,a*200+b], '-k', label='Linear Fit')
 
 # data from http://www.barnardmicrosystems.com/UAV/engines/2_stroke.html
 a_ref = 76.0
 b_ref = 502.0
-plt.plot([0,200], [b_ref,a_ref*200+b_ref], '-k')
+plt.plot([0,200], [b_ref,a_ref*200+b_ref], '--k', label='Barnard Microsystems Fit')
 
 # quadratic model
 A = np.array([np.multiply(powerdata[:,0],powerdata[:,0]), powerdata[:,0]])
@@ -63,7 +70,8 @@ a, b  = np.linalg.lstsq(A.T,f)[0]
 print a,b
 x = np.linspace(0,200, num=100)
 y = a*np.multiply(x,x) + b*x
-plt.plot(x, y, '-k')
+plt.plot(x, y, '-.k', label='Quadratic Fit thru Origin')
[…]

### 2016-04-10 matlab/thrust.m
working thrust model

@@ -12,7 +12,7 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
     f_rate = x(6)/10^3;
 
     V_c = 0;                % Vertical velocity is zero because of hover.
-    a = .1091;               % Lift curve slope for naca 0012
+    a = .1091*180/pi;               % Lift curve slope for naca 0012
     rho = 1.225;            % Air density, kg/m^3
     C = .09*R;              % Chord length, based on rc rotor blades (extrapolated).
     T = 0;
@@ -31,8 +31,8 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
     for i=1:m
         dr = R/m;
         r = i*R/m-.5*R/m;
-        s = n*C/pi*r;
-        phi = 1/16*(-sqrt(a)*sqrt(s)*sqrt(a*s+32*theta)-a*s); %atan(U_P/U_T);          % Small angle approximation because U_P << U_T.
+        s = N*C/(pi*r);
+        phi = 1/16*(sqrt(a)*sqrt(s)*sqrt(a*s+32*theta)-a*s); %atan(U_P/U_T);          % Small angle approximation because U_P << U_T.
         alpha = theta - phi;
         C_l = -98.031*alpha^3 + 0.9481*alpha^2 + 8.2078*alpha - 0.009;          % Lift coefficient based on Xfoil data.
         C_d = 51.031*alpha^4 - 0.9405*alpha^3 - 0.5316*alpha^2 + 0.0191*alpha + 0.0168;   % Drag coefficient based on xfoil data.
@@ -48,18 +48,20 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
         dT = .5*rho*a*N*omega^2.*C.*(theta-phi).*r.^2.*dr;
 %         dQ = N*(L*phi+D)*r*dr;
 %         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr;
-        dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr;
+        dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr*N;
+        
         dP = dQ*omega;
         T = dT*n + T;
         Q = dQ*n + Q;
         P = dP*n + P; 
     end
     
-    T_produced = T*n;
-    power_required = P*n;
+    T_produced = T;
+    power_required = P;
     power_produced = f_rate*E_rhof*n_engine;
[…]

### 2016-04-11 matlab/thrust.m
Deleted commented code

@@ -36,18 +36,8 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
         alpha = theta - phi;
         C_l = -98.031*alpha^3 + 0.9481*alpha^2 + 8.2078*alpha - 0.009;          % Lift coefficient based on Xfoil data.
         C_d = 51.031*alpha^4 - 0.9405*alpha^3 - 0.5316*alpha^2 + 0.0191*alpha + 0.0168;   % Drag coefficient based on xfoil data.
-        t_c = C_l/6;
-        lambda_i = sqrt(s*t_c/2);
-        v_i = omega*r*lambda_i;
-        U_T = omega*r;          % Tangential component of velocity.
-        U_P = V_c + v_i;          % Perpendicular component of velocity.
-  
-        L = .5*rho*C*a*(U_T^2*theta-U_P*U_T);
-        D = .5*rho*U_T^2*C*C_d;
-%         dT = N*L*dr;
+       
         dT = .5*rho*a*N*omega^2.*C.*(theta-phi).*r.^2.*dr;
-%         dQ = N*(L*phi+D)*r*dr;
-%         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr;
         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr*N;
         
         dP = dQ*omega;
@@ -62,14 +52,7 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
     T_required     = AUW*9.8;   % 1.5 multiplier for controllability.
     FT = -f_cap/f_rate;             % multiply by negative to allow for minimization.
     AUW_limit = 24.9*9.8;   % 55 lbs
-%     persistent iter;
-%     if isempty(iter)
-%         iter = 0;
-%     end
-%     iter = iter +1;
-%     if iter > 1000
-%         keyboard
-%     end
+
 end
 
 

### 2016-04-11 matlab/autodiff/@adiff/plus.m
Added all matlab code

@@ -1,0 +1,27 @@
+function c = plus(a,b)
+% PLUS implements a+b, where either a or b is an adiff object.
+
+switch [class(a),class(b)]
+   
+case 'adiffdouble'
+   c = adiff(a.x+b, repmat(a.dx,length(b),1), a.root);
+   
+case 'doubleadiff'
+   c = adiff(b.x+a, repmat(b.dx,length(a),1), b.root);
+   
+case 'adiffadiff'
+   checkroot(a,b);
+   if size(a.dx,1)~=size(b.dx,1)
+      if size(a.dx,1)==1
+         a.dx = repmat(a.dx,size(b.dx,1),1);
+      elseif size(b.dx,1)==1
+         b.dx = repmat(b.dx,size(a.dx,1),1);
+      end
+   end
+   c = adiff( a.x+b.x,a.dx+b.dx, a.root);
+   
+otherwise
+   error(['Can''t add ',class(a),' and ',class(b)]);
+   
+end
+

### 2016-04-11 matlab/autodiff/interp1.m

@@ -1,0 +1,24 @@
+function yi = interp1(x,y,xi,method)
+% interp1(x,y,xi) does interpolation for autodiff objects 
+%   - note that x must be a double
+
+if nargin==3, method='linear'; end
+switch lower(method)
+case {'linear','*linear'}
+   x  = x(:);
+   [x,xidx] = sort(x); % which means that x can't be an adiff object
+   y  = y(xidx);
+   dy = diff(y)./diff(x);
+   xidx = zeros(size(xi));
+   for i=1:length(xi)
+      if xi(i)==x(end)
+         xidx(i) = length(x)-1;
+      else
+         xidx(i) = sum(x<=xi(i));
+      end
+   end
+   yi = y(xidx)+dy(xidx).*(xi(:)-x(xidx));
+   
+otherwise
+   error([method,' method is not supported for autodiff objects']);
+end

### 2016-04-11 matlab/grad.m

@@ -1,0 +1,10 @@
+function [g] = grad(x0)
+
+addpath(strcat(pwd,'/autodiff'));
+
+x = adiff(x0);
+[a,~,~,~,~] = thrust(x);
+[~,g] = adiffget(a);
+
+
+end

### 2016-04-11 matlab/initializer.m

@@ -1,0 +1,19 @@
+clear all;
+clc;
+clf;
+close all;
+
+
+
+tic
+[opt_A,fval,exitflag,output_fmin,lambda,gradient,constraints] = optimizer();
+toc
+
+Num_blades = opt_A(1);
+rotor_radius = opt_A(2)/10;
+omega = opt_A(3)/2/pi*60*100;
+Num_rotors = opt_A(4);
+Blade_pitch = opt_A(5)*180/pi/10;
+fuel_rate = opt_A(6)/10^3;
+Flight_time = -fval*1e4/3600;
+

### 2016-04-11 matlab/optimizer.m

@@ -1,0 +1,59 @@
+function [opt_A,fval,exitflag,output_fmin,lambda,gradient,constraints] = optimizer()
+
+
+
+options = optimoptions('fmincon',...
+        'Algorithm', 'sqp', ...  % choose one of: 'interior-point', 'sqp', 'active-set', 'trust-region-reflective'
+        'AlwaysHonorConstraints', 'bounds', ...  % forces interior point algorithm to always honor bounds
+        'display', 'iter-detailed', ...  % display more information
+        'Tolx', 1e-15, ...
+        'MaxIter', 10000, ...  % maximum number of iterations
+        'MaxFunEvals', 100000, ...  % maximum number of function calls
+        'TolCon', 1e-7, ...  % convergence tolerance on constraints
+        'TolFun', 1e-7, ...  % convergence tolerance on function value
+        'GradObj', 'on', ...  % supply gradients of objective
+        'GradConstr', 'off', ...  % supply gradients of constraints
+        'Diagnostics', 'on', ...  % display diagnotic information
+        'PlotFcns',{@optimplotfval,@optimplotfirstorderopt});
+
+A0 = [2,.5*10,2*pi*2000/60/100,4,5*pi/180*10,.001*10^3]; % Initial guess on number of blades in each rotor, 
+                                       %rotor radius, angular velocity, number of rotors, 
+                                       %and blade pitch, fuel consumption rate.                
+
+Aeq = [];
+Beq = [];
+lb = [2, .01*10, 2*pi*100/60/100, 1, 0, 0.0001848346*10^3]; 
+ub = [2, .7*10, 2*pi*4500/60/100, 4, 20*pi/180*10, 0.0246446*10^3];
+
+
+[opt_A,fval,exitflag,output_fmin,lambda,gradient] = fmincon(@obj, A0, [], [], Aeq, Beq, lb, ub,@con,options);
+
+[~,constraints.power_produced, constraints.power_required, constraints.thrust_produced, ...
+    constraints.thrust_required] = thrust(opt_A);
+
+
+%     ---------- Objective Function ------------------
+    function [J,g] = obj(x)
+        
+        [f,~,~,~,~] = thrust(x);
+        dfdx = grad(x);
+        J = f / 1e4;  % scale to make objective of order(1)
+        g = dfdx / 1;
+    end
+
+
+
+    function [cin, ceq, dcin, dceq] = con(x)
+        
[…]

### 2016-04-11 presentation/presentation.tex
added an outline to the presentation
Added current capabilities

@@
 % presentation info
-\title[Short Title]{GAS Quad}
-\author{Smart Person}
+\title{Multi-dimensional Optimization of Gasoline-fueled Variable Pitch Multirotor Aircraft}
+%\title[Short Title]{GAS Quad}
+\author{Dallin Briggs, Gary Ellingson}
 \institute[MAGICC Lab]{\includegraphics[height=0.5in]{logo_gray}}
@@
 
+\begin{frame}{Outline}
+\begin{itemize}
+	\item{Motivation and Background}
+	\item{---figure of multirotor trends in flight time}
+	\item{---focus on OTS stuff bases our design on empirical models}
+
+	\item{Theory - models}
+	\item{---Engine model}
+	\item{---Aero model}
+
+	\item{Optimization}
+	\item{---Objective only directly a function of 2 design variables}
+	\item{---constraints}
+
+	\item{Results}
+	\item{lessons learned}
+	\item{questions}
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Motivation and Background}
+	\begin{center}
+	Currently available multirotor capabilities
+	\end{center}
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.70\textwidth]{current_capabilities.png}
+			\label{current_capabilities}
+		\end{center}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Models}
+	\begin{itemize}
+		\item{To represent the aerodynamics of each rotor, we used Combined Differential Blade Element and Momentum Theory for nonuniform inflow.}
+		\begin{center}
+		$ dT = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr $
+
+		$ dQ = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr $
+		\end{center}
+	\end{itemize}
+
+\end{frame}
+
+\begin{frame}{Optimization Method}
+
+\end{frame}
+
 \end{document}

### 2016-04-11 python/project.py
moved some figures around

@@ -57,8 +57,8 @@ if __name__ == '__main__':
 
     plt.figure(2)
     plt.xlabel('Useful Payload (kg)')  # labels for axes
-    plt.ylabel('Engine size (cc)')
-    plt.title('Verying Engine Size with payload')
+    plt.ylabel('Engine Size (cc)')
+    plt.title('Engine Size with Payload')
     plt.plot(payloads, engine_size, '-k', label='Data')
     # plt.legend(loc='upper left')
     plt.show()

### 2016-04-11 presentation/presentation.tex
Added blank slides
added engine models stuff to the presentation
Updated questions slide
moved some figures around

@@
 
+\begin{frame}{Introduction}
+
+	What is a gas quad...
+\end{frame}
+
 \begin{frame}{Motivation and Background}
@@
 		\begin{center}
-			\includegraphics[width=.70\textwidth]{current_capabilities.png}
+			\includegraphics[width=.70\textwidth]{../current_capabilities.png}
 			\label{current_capabilities}
@@
 	\end{figure}
+\end{frame}
+
+\begin{frame}{Engine Models}
+
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.70\textwidth]{../mass.png}
+			\label{fig:eng_mass}
+		\end{center}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Engine Models}	
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.70\textwidth]{../max_power.png}
+			\label{fig:eng_power}
+		\end{center}
+	\end{figure}
+
 \end{frame}
@@
 		\begin{center}
-		$ dT = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr $
-
-		$ dQ = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr $
+		$$ dT = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr $$
+		$$ dQ = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr $$
 		\end{center}
@@
 \begin{frame}{Optimization Method}
+	Design Variables
+	\begin{itemize}
+		\item{Engine Size}
+		\item{Rotor Size}
+		\item{Fuel Capacity}
+		\item{Fuel Burn Rate}
+		\item{theta hover}
+		\item{omega}
+	\end{itemize}
+
+	Flight Time = f(Fuel Capacity/Fuel Rate)
+
+	C = ...
+
+	SNOPT
+\end{frame}
+
+\begin{frame}{Results}
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.70\textwidth]{../pareto_front.png}
+			\label{fig:pareto}
+		\end{center}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Results}
+	\begin{figure}
+		\begin{center}
+			\includegraphics[width=.70\textwidth]{../engine_size_vs_payload.png}
+			\label{fig:pareto}
+		\end{center}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Lessons Learned}
+	\begin{itemize}
+		\item{Helicopter rotor dynamics are difficult to conceptualize and calculate.}
+		\item{Over-simplification of aerodynamic model leads to unrealistic results.}
+		\item{large, slow props - https://youtu.be/syJq10EQkog?t=7s}
+
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Questions}
+
+		\begin{center}
+			Questions?
+		\end{center}
 

### 2016-04-11 python/flight_time.py
added the multipayload optimization stuff

@@ -106,7 +106,7 @@ def scale(x):
     eng_displace = eng_displace/80.0
     fuel_rate = fuel_rate/.02
     omega_r = omega_r/1500.0
-    R_rotor = R_rotor/0.705
+    R_rotor = R_rotor
     fuel_cap = fuel_cap
     theta_hover = theta_hover/20
 
@@ -118,7 +118,7 @@ def unscale(x):
     eng_displace = eng_displace*80.0
     fuel_rate = fuel_rate*.02
     omega_r = omega_r*1500.0
-    R_rotor = R_rotor*0.705
+    R_rotor = R_rotor
     fuel_cap = fuel_cap
     theta_hover = theta_hover*20
 
@@ -169,7 +169,7 @@ def simple_aero_model(AUW, omega, R_rotor, c_rotor, theta_hover, rho_air, n_roto
     T = n_rotors*2*T
     Q = n_rotors*2*Q
 
-    return T, 8*Q
+    return T, 4*Q
 
 def other_aero_model(omega, R_rotor, c_rotor, theta, rho_air, n_rotors):
 

### 2016-04-11 python/project.py

@@ -19,7 +19,7 @@ if __name__ == '__main__':
     theta_hover = 20 # deg
     x0 = scale([eng_displace, fuel_rate, omega_r, R_rotor, fuel_cap, theta_hover])
     # x0 = [ 4.3441622,   0.16620041,  0.10885692,  2.78460696,  6.76453586, 0.85782237]
-    x0 = [ 4.61280557,  0.17376677,  0.08200489,  3.4691025,   7.32432873, 0.86800068]
+    x0 = [ 4.61280557,  0.17376677,  0.08200489,  1.5,   7.32432873, 0.86800068]
 
     lb = [0, 0, 0, 0, 0, 0]
     ub = scale([1200.0, 1.0, 7500.0, 3.0, 12.0, 30.0])
@@ -36,7 +36,10 @@ if __name__ == '__main__':
 
     out = obj_func_print(xopt)
 
-    steps = 30
+
+    # Create the pareto frunt
+    '''
+    steps = 10
     payloads = np.linspace(0.0, 8.0, num=steps)
     flight_times = np.zeros(steps)
     engine_size = np.zeros(steps)
@@ -46,6 +49,7 @@ if __name__ == '__main__':
         xopt, fopt, info = optimize(obj_func, x0, lb, ub, optimizer)
         flight_times[i] = -fopt/60
         engine_size[i] = xopt[0]*80.0
+        engine_size[i] = xopt[3]
 
     plt.figure()
     plt.xlabel('Useful Payload (kg)')  # labels for axes
@@ -62,3 +66,31 @@ if __name__ == '__main__':
     plt.plot(payloads, engine_size, '-k', label='Data')
     # plt.legend(loc='upper left')
     plt.show()
+    '''
+
+    # integrate over multiple payloads
+    def multiOpt(x):
+        ft = 0
+        c = []
+        setPayload(5.0);
+        fti, ci = obj_func(x)
+        ft += fti
+        c = ci
+        setPayload(7.0);
+        fti, ci = obj_func(x)
+        ft += fti
+        c = np.concatenate((c, ci), axis=0)
+        setPayload(3.0);
+        fti, ci = obj_func(x)
+        ft += fti
+        c = np.concatenate((c, ci), axis=0)
+        return ft/3.0, c
+
+    print "-----"
+
+
+    xopt, fopt, info = optimize(multiOpt, xopt, lb, ub, optimizer)
+    print 'SNOPT:', xopt, fopt, info
+    print multiOpt(xopt)
+
+    out = obj_func_print(xopt)

### 2016-04-11 presentation/presentation.tex
added the multipayload optimization stuff
Merged conflicts

@@
 
-\begin{frame}{Presentation Guidelines}
-
-You will sign up for a 10 minute presentation time slot.
-The presentation is graded only for completion.
-Group projects will have 8 minutes to present and 2 minutes for Q\&A.
-Individuals will have 6 minutes to present and 2 minutes for Q\&A.
-Obviously you will not have a lot of time to spend on details.
-Instead, you should focus on motivation (what the problem is and why it is important), the solution approach, and the most important results.
-Each member of the team should have a role in the presentation.
-
-\end{frame}
-
-\begin{frame}{Outline}
-\begin{itemize}
-	\item{Motivation and Background}
-	\item{---figure of multirotor trends in flight time}
-	\item{---focus on OTS stuff bases our design on empirical models}
-
-	\item{Theory - models}
-	\item{---Engine model}
-	\item{---Aero model}
-
-	\item{Optimization}
-	\item{---Objective only directly a function of 2 design variables}
-	\item{---constraints}
-
-	\item{Results}
-	\item{lessons learned}
-	\item{questions}
-\end{itemize}
-\end{frame}
-
 \begin{frame}{Introduction}
 
-	What is a gas quad...
+	What is a gas multirotor?
+	\begin{itemize}
+	\item{Uses a single small two-stroke engine.}
+	\item{Power is transfered through torque tubes to the rotors.}
+	\item{All rotors must be variable pitch.}
+	\item{Theoretically capable of longer flight times than battery powered multirotors because of higher specific energy of gasoline.}
+	\end{itemize}
 \end{frame}
@@
 
-\begin{frame}{Models}
+\begin{frame}{Aerodynamic Model}
 	\begin{itemize}
@@
 		\item{Fuel Burn Rate}
-		\item{theta hover}
-		\item{omega}
+		\item{Rotor Pitch Angle}
+		\item{Head Speed}
 	\end{itemize}
 
-	Flight Time = f(Fuel Capacity/Fuel Rate)
+	$Flight Time = f(Fuel Capacity, Fuel Rate)$
 
-	C = ...
+	$C: P_{Shaft} \geq P_{Rotor}, T \geq W_{total}, P_{max} \geq P_{Shaft}, 55lbs \geq W_{total} $
 
-	SNOPT
+	SNOPT to optimize
+
 \end{frame}
@@
 
+\begin{frame}{Results}
+	Simultaneously optimize over 3, 5, and 7 kg payload
+	\begin{itemize}
+		\item{Engine Size: 177 cc} 
+		\item{Rotor Size: 2.16 m}
+		\item{Fuel Capacity: 6.45 kg}
+		\item{Fuel Burn Rate: 0.0019 kg/s}
+		\item{Rotor Pitch Angle: 17.618 deg}
+		\item{Head Speed: 135.09 rpm}
+	\end{itemize}
+
+	Flight Time > 42.47 minutes
+
+	Payload < 7 kg
+\end{frame}
+
 \begin{frame}{Lessons Learned}
 	\begin{itemize}
+		\item{Large, Slow props - https://youtu.be/syJq10EQkog?t=7s}
 		\item{Helicopter rotor dynamics are difficult to conceptualize and calculate.}
 		\item{Over-simplification of aerodynamic model leads to unrealistic results.}
-		\item{large, slow props - https://youtu.be/syJq10EQkog?t=7s}
-
+		\item{Scaling the problem was critical}	
 	\end{itemize}

### 2016-04-11 python/project.py
small revisions

@@ -38,9 +38,9 @@ if __name__ == '__main__':
 
 
     # Create the pareto frunt
-    '''
-    steps = 10
-    payloads = np.linspace(0.0, 8.0, num=steps)
+
+    steps = 20
+    payloads = np.linspace(0.0, 10.0, num=steps)
     flight_times = np.zeros(steps)
     engine_size = np.zeros(steps)
     for i in range(steps):
@@ -49,7 +49,7 @@ if __name__ == '__main__':
         xopt, fopt, info = optimize(obj_func, x0, lb, ub, optimizer)
         flight_times[i] = -fopt/60
         engine_size[i] = xopt[0]*80.0
-        engine_size[i] = xopt[3]
+        # engine_size[i] = xopt[3]
 
     plt.figure()
     plt.xlabel('Useful Payload (kg)')  # labels for axes
@@ -94,3 +94,4 @@ if __name__ == '__main__':
     print multiOpt(xopt)
 
     out = obj_func_print(xopt)
+    '''

### 2016-04-12 matlab/initializer.m
Added engine/frame models to matlab aerodynamics

@@ -15,5 +15,6 @@ omega = opt_A(3)/2/pi*60*100;
 Num_rotors = opt_A(4);
 Blade_pitch = opt_A(5)*180/pi/10;
 fuel_rate = opt_A(6)/10^3;
+engine_size = opt_A(7)*100;
 Flight_time = -fval*1e4/3600;
 

### 2016-04-12 matlab/optimizer.m

@@ -3,7 +3,7 @@ function [opt_A,fval,exitflag,output_fmin,lambda,gradient,constraints] = optimiz
 
 
 options = optimoptions('fmincon',...
-        'Algorithm', 'sqp', ...  % choose one of: 'interior-point', 'sqp', 'active-set', 'trust-region-reflective'
+        'Algorithm', 'interior-point', ...  % choose one of: 'interior-point', 'sqp', 'active-set', 'trust-region-reflective'
         'AlwaysHonorConstraints', 'bounds', ...  % forces interior point algorithm to always honor bounds
         'display', 'iter-detailed', ...  % display more information
         'Tolx', 1e-15, ...
@@ -16,26 +16,27 @@ options = optimoptions('fmincon',...
         'Diagnostics', 'on', ...  % display diagnotic information
         'PlotFcns',{@optimplotfval,@optimplotfirstorderopt});
 
-A0 = [2,.5*10,2*pi*2000/60/100,4,5*pi/180*10,.001*10^3]; % Initial guess on number of blades in each rotor, 
+A0 = [2,.5*10,2*pi*2000/60/100,4,5*pi/180*10,.001*10^3,100/100]; % Initial guess on number of blades in each rotor, 
                                        %rotor radius, angular velocity, number of rotors, 
                                        %and blade pitch, fuel consumption rate.                
 
 Aeq = [];
 Beq = [];
-lb = [2, .01*10, 2*pi*100/60/100, 1, 0, 0.0001848346*10^3]; 
-ub = [2, .7*10, 2*pi*4500/60/100, 4, 20*pi/180*10, 0.0246446*10^3];
+lb = [2, .01*10, 2*pi*00/60/100, 1, 0, 0.000001848346*10^3,0]; 
+ub = [2, 10*10, 2*pi*4500/60/100, 4, 20*pi/180*10, 0.0246446*10^3,500/100];
 
 
 [opt_A,fval,exitflag,output_fmin,lambda,gradient] = fmincon(@obj, A0, [], [], Aeq, Beq, lb, ub,@con,options);
 
 [~,constraints.power_produced, constraints.power_required, constraints.thrust_produced, ...
-    constraints.thrust_required] = thrust(opt_A);
+    constraints.thrust_required,constraints.AUW,constraints.AUW_limit, ...
+    constratins.engine_power] = thrust(opt_A);
 
 
 %     ---------- Objective Function ------------------
     function [J,g] = obj(x)
         
[…]

### 2016-04-12 matlab/thrust.m

@@ -1,4 +1,4 @@
-function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
+function [FT,power_produced,power_required,T_produced,T_required,AUW,AUW_limit,engine_power] = thrust(x)
 
     % This method of calculating thrust uses blade element theory
     %x consists of the number of blades in each rotor, rotor radius, 
@@ -10,22 +10,24 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
     n = x(4);
     theta = x(5)/10;
     f_rate = x(6)/10^3;
+    engine_size = x(7)*100;
+    
 
-    V_c = 0;                % Vertical velocity is zero because of hover.
+    
     a = .1091*180/pi;               % Lift curve slope for naca 0012
     rho = 1.225;            % Air density, kg/m^3
     C = .09*R;              % Chord length, based on rc rotor blades (extrapolated).
     T = 0;
     Q = 0;
     P = 0;
-    n_engine = 0.15;                            % Efficiency of the engine
-    E_rhof = 32.4e6;                            %Energy density of fuel (Joules/L)
-    f_cap = 7;             % In liters.
-    W_motor = 2.1;                              %Motor weight
+    n_engine = 0.1;                            % Efficiency of the engine
+    E_rhof = 44.4e6;                            %Energy density of fuel (Joules/L)
+    f_cap = 7;                                  % In liters.
+    W_motor = 0.02144822774*engine_size + 0.254117290343;                              %Motor weight
     W_f = f_cap*.75;                            %Fuel weight (kg)
     payload = 5;                                %Payload weight
-    W_frame = 7;                                %Frame weight
-    AUW = W_motor + W_f + W_frame + payload;    %All up weight in kg (about 52 lbs)
+    W_frame = 4.26*R;                                %Frame weight
+    AUW = (W_motor + W_f + W_frame + payload)*9.8;    %All up weight in kg (about 52 lbs)
     
     m = 10;
     for i=1:m
@@ -40,18 +42,19 @@ function [FT,power_produced,power_required,T_produced,T_required] = thrust(x)
         dT = .5*rho*a*N*omega^2.*C.*(theta-phi).*r.^2.*dr;
         dQ = .5*rho*omega^2*r^3*C*(C_d+phi*C_l)*dr*N;
         
-        dP = dQ*omega;
+%         dP = dQ*omega;
[…]

### 2016-04-12 matlab/initializer.m
Added fuel capacity as design variable

@@ -17,4 +17,5 @@ Blade_pitch = opt_A(5)*180/pi/10;
 fuel_rate = opt_A(6)/10^3;
 engine_size = opt_A(7)*100;
 Flight_time = -fval*1e4/3600;
+fuel_capacity = opt_A(8);
 

### 2016-04-12 matlab/optimizer.m

@@ -16,14 +16,14 @@ options = optimoptions('fmincon',...
         'Diagnostics', 'on', ...  % display diagnotic information
         'PlotFcns',{@optimplotfval,@optimplotfirstorderopt});
 
-A0 = [2,.5*10,2*pi*2000/60/100,4,5*pi/180*10,.001*10^3,100/100]; % Initial guess on number of blades in each rotor, 
+A0 = [2,.5*10,2*pi*2000/60/100,4,5*pi/180*10,.001*10^3,100/100,5]; % Initial guess on number of blades in each rotor, 
                                        %rotor radius, angular velocity, number of rotors, 
                                        %and blade pitch, fuel consumption rate.                
 
 Aeq = [];
 Beq = [];
-lb = [2, .01*10, 2*pi*00/60/100, 1, 0, 0.000001848346*10^3,0]; 
-ub = [2, 10*10, 2*pi*4500/60/100, 4, 20*pi/180*10, 0.0246446*10^3,500/100];
+lb = [2, .01*10, 2*pi*00/60/100, 1, 0, 0.00001848346*10^3,0,0]; 
+ub = [2, 10*10, 2*pi*4500/60/100, 4, 22*pi/180*10, 0.0246446*10^3,500/100,30];
 
 
 [opt_A,fval,exitflag,output_fmin,lambda,gradient] = fmincon(@obj, A0, [], [], Aeq, Beq, lb, ub,@con,options);

### 2016-04-12 matlab/thrust.m

@@ -11,7 +11,7 @@ function [FT,power_produced,power_required,T_produced,T_required,AUW,AUW_limit,e
     theta = x(5)/10;
     f_rate = x(6)/10^3;
     engine_size = x(7)*100;
-    
+    f_cap = x(8);                                  % In liters.
 
     
     a = .1091*180/pi;               % Lift curve slope for naca 0012
@@ -22,7 +22,7 @@ function [FT,power_produced,power_required,T_produced,T_required,AUW,AUW_limit,e
     P = 0;
     n_engine = 0.1;                            % Efficiency of the engine
     E_rhof = 44.4e6;                            %Energy density of fuel (Joules/L)
-    f_cap = 7;                                  % In liters.
+    
     W_motor = 0.02144822774*engine_size + 0.254117290343;                              %Motor weight
     W_f = f_cap*.75;                            %Fuel weight (kg)
     payload = 5;                                %Payload weight

### 2016-04-12 paper.tex
updated some figures
added the pareto front payload vs flight time
moved some figures around
Updated aero model

@@
 
-All commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
+Nearly all commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
 Although these batteries are very good and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
@@
 	\begin{center}
-		\includegraphics[width=.40\textwidth]{current_capabilities.jpg}
+		\includegraphics[width=.40\textwidth]{current_capabilities.png}
 		\caption{\textit{Graphic showing current platforms and how our would be better.}}
@@
 
+\subsection{Aerodynamics}
+
 Much of the work done in the methodology of optimizing flight time was to find adequate models for the thrust, torque, and power required from the main rotor disks.
+Since each rotor will be designed using small-scale helicopter components that consist of variable-pitch rotors that have no twist, a simplified helicopter rotor aerodynamic model should be sufficient for these calculations.
+
+Many of the existing aerodynamic models used in the aerospace industry to represent helicopter rotors are extremely complex and require specialized software tools (many of which are proprietary).
+High fidelity aerodynamic models generally require lots of computational power; so in order to focus on optimization methods rather than the complex aerodynamics of the system, several simplifications were made.
+
+Combined differential blade element and momentum theory for nonuniform inflow was used to calculate the thrust and the torque differentially across each rotor blade \cite{bramwell2001bramwell}.
+The differential thrust is given by
+
+\begin{equation}
+	dT = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr,
+	\label{thrust_eqn}
+\end{equation}
+
+where $\rho$ is the air density, \textit{a} is the lift-curve slope for the blade airfoil, $\Omega$ is the angular velocity of the rotors, \textit{r} is the radius of the annulus of the rotor being differentiated, $\theta$ is the blade pitch angle measured from the horizontal plane formed by the rotational disc of rotor blades, \textit{c} is the chord of the rotor, and $\phi = \tan^{-1}(U_p/U_T) $ is the inflow angle at the blade element.
+$U_P$ and $U_T$ are the components of air velocity relative to blade element perpendicular and tangential to the plane of rotation, respectively.
+The total thrust on each rotor is calculated by integrating equation \ref{thrust_eqn} across the rotor radius, \textit{R}.
+The total thrust for the multirotor can then be calculated by multiplying the blade thrust by the number of \textit{N} rotors.
+
+The differential torque for each rotor blade is given by
+
+\begin{equation}
+	dQ = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr,
+	\label{torque_eqn}
+\end{equation}
+
+where $\delta$ is the local blade drag coefficient, and $C_L$ is the lift coefficient. The total torque for each blade and the multirotor can be found using the same method for calculating the thrust.
+
+The differential power produced by each rotor is given by
+
+\begin{equation}
+	dP = dQ\Omega,
+\end{equation}
+
+and the total power can be calculated using the same method as thrust and torque.
 
@@
 \begin{figurehere}
-	\includegraphics[width=0.5\textwidth]{payload.jpg}
+	\includegraphics[width=0.5\textwidth]{pareto_front.png}
 	\caption{Stand in figure showing optimal flight time as a function of payload.}

### 2016-04-12 paper.tex
Removed old aero stuff

@@
 
-First model used in the beginning of the project used a simplified version of Momentum theory and actuator disk theory from \cite{helicopters2016}.
-This model took into account the number of rotors, the rotor radius, the disc loading of each rotor, and how much fuel was available onboard the aircraft.
-However, this model did not take into account the blade speed, blade pitch angle, or the number of blades per rotor.
-The model did, however, give us a baseline for initial optimization results.
-
-Second model used to develop the optimization framework for this project was based on an MIT lecture given on aircraft propeller performance given in \cite{mit2016}.
-In order to test the theory of this model and also to further the progress of the optimization model, only two of the design variables were chosen to optimize: rotor radius and fuel consumption rate, $f_{rate}$.
-
-To properly analyses the system, power required ($P_{req}$) and thrust produced ($T_{prod}$) have to be calculated to make sure that the rotor radius had a relationship to the fuel consumption rate.
-These calculations used some approximations in the rotor dynamics.
-The following calculations come from a MIT lecture on thermodynamics.
-
-In order to calculate thrust of each rotor blade, the thrust has to be calculated differentially along the blade because of the angular velocity. The differential thrust is given by 
-\begin{equation}
-	dT = dLcos(\phi + \alpha_i) - dDsin(\phi + \alpha_i)
-\end{equation}
-
-where $\phi = \tan^{-1}(\frac{V_i}{V_n})$, and $\alpha_i = \theta - \tan^{-1}(\frac{V_i}{V_n})$ and $\theta$ is the blade pitch angle (held constant in this case of the optimization).
-$dL$ and $dD$ are the differentials of the lift and drag, respectively, and are given by
-\begin{equation}
-	dL = \frac{1}{2}\rho V_e^2 c C_l dr
-\end{equation}
-
-and
-
-\begin{equation}
-	dD = \frac{1}{2}\rho V_e^2 c C_d dr.
-\end{equation}
-
-The coefficients of lift and drag were held constant at $C_l = 0.5438$ and $C_d = 0.0141$ (obtained from a symmetric airfoil pitch at $\alpha = 5$ degrees).
-Splitting the rotor into 10 sections and integrating seemed to give a reasonable estimate to the thrust produced by each rotor.
-This was then constrained by making sure the thrust produced was greater than or equal to the thrust required, which was
-\begin{equation}
-	T_{req} = AUW\times g.
-\end{equation}
-
-In order to make sure that enough power was being produced by the engine (and prevent the fuel consumption rate to going to 0), the power produced was given by 
-\begin{equation}
-	P_{prod} = f_{rate}E_{\rho f} \eta_{engine}
-\end{equation}
-
-and was made sure that it was greater than or equal to the power required. To calculate the power required, the torque required to turn each rotor was calculated differentially by the equation
-\begin{equation}
-	dQ = r[dLsin(\phi + \alpha_i) + dDcos(\phi + \alpha_i)].
-\end{equation}
-
-The power required was then calculated by 
-\begin{equation}
-	P_{req} = \omega_rQ_{req}2n_{rotors}.
-\end{equation}
-
-This model did converge to an optimal solution when executed with gradient based optimization methods discussed in the next section.
-However, in this model, not all of the design variables were implemented due to the complexity these additional variables brought to the calculus in this method.
-
-The third model used to represent the aerodynamics and thrust of each rotor blade was from \cite{chen1979simplified}, a NASA model used in helicopter simulations several years ago.
-Instead of trying to implement blade element theory across the blade, this new model is able to approximate the thrust and torque required for the rotor given numerous variables.
-Instead of only varying rotor radius and fuel consumption rate, like last homework's model, this model attempts to optimize given rotor radius, fuel consumption rate, number of blade per rotor, number of rotors, rotor speed, and rotor pitch.
-The thrust, T, is given by equation \ref{thrust} and the torque, Q, is represented by equations \ref{torque} and \ref{torquecoeff}.
-
-\begin{multline}
-T = \frac{N}{2} \rho a c R (\omega R)^2 (.5 (1-\epsilon^2) \lambda \\
- + \theta_0 (\frac{1}{3} + \frac{\mu^2}{2} (1-\epsilon)) + \theta_t (.25+\frac{\mu^2}{2} (1-\epsilon^2))\\
-- \frac{\mu}{2} (1-\epsilon^2) (B_{lc}-K_l b_l) - a_0 (\frac{1}{3} +\frac{\mu^2}{2} (1-\epsilon)) K_l \\
-+ a_l (\frac{\mu}{2} \epsilon (1-\epsilon)) - \frac{\dot{a_0}}{\omega} (\frac{1}{3} -\frac{\epsilon}{2}) + \frac{\dot{b_l}}{\omega} (\frac{\mu}{4} (1-\epsilon)^2) \\
-+ \frac{\mu}{4} (1-\epsilon^2) (\frac{p}{\omega cos(\beta_w)} + \frac{q}{\omega sin(\beta_w))} ) - N M_\beta \ddot{a_0}
-\label{thrust}
-\end{multline}
-and
-\begin{equation}
-Q = N/2 \rho a c R^2 (\omega R)^2 (Q_{coeff})
-\label{torque}
-\end{equation}
-where
-\begin{multline}
-Q_{coeff} = \delta/(4 a) (1 + (1-\epsilon^2) \mu^2) \\
- - (\theta_0 - K_l a_0)(\lambda/3+(\epsilon/3-.25) \dot{a_0}/\omega + \mu/6 \dot{b_l}/\omega\\  
-- \mu \epsilon/4 (\dot{b_l}/\omega-a_l) + \mu/6 (p/\omega cos(\beta_w) +q/\omega sin(\beta_w))) \\
-- (A_{lc}-K_l a_l) ((.125-\epsilon/6) (\dot{a_l}/\omega+b_l)-\mu/6 a_0 \\
-+b_l/16 (1-\epsilon^2) \mu^2+1/8 (-p/\omega sin(\beta_w)+q/\omega cos(\beta_w)))\\ 
-- (B_lc-K_l b_l) ((.125-\epsilon/6) (\dot{b_l}/\omega-a_l)\\
-+(\epsilon/4-1/6) \mu \dot{a_0}/\omega+.5 (1-\epsilon^2) (\mu \lambda/2+a_l/8 \mu^2)\\ 
-+ 1/8 (p/\omega cos(\beta_w)+q/\omega sin(\beta_w))) - \theta_t (\lambda/4\\ 
-+ (\epsilon/4-1/5) \dot{a_0}/\omega+\mu/8 (\dot{b_l}/\omega)-\epsilon \mu/6 (\dot{b_l}/\omega-a_l))\\  
-- .5 (1-\epsilon^2) (\lambda^2+\lambda \mu a_l+2 \lambda \epsilon \dot{a_0}/\omega\\  
-+\mu \epsilon (a_l \dot{a_0}/\omega+a_0 (\dot{a_l}/\omega+b_l))\\  
-+ \mu^2 (a_0^2/2+3/8 a_l^2+1/8 b_l^2)) + \mu/3 (a_l (\dot{a_0}/\omega) \\
-+ a_0 (\dot{a_l}/\omega+b_l))+2/3 \lambda (\dot{a_l}/\omega)\\  
-- (-\mu/3 a_0 + (.25-\epsilon/3) (\dot{a_l}/\omega+b_l)) \\ 
-(-p/\omega sin(\beta_w)+q/\omega cos(\beta_w))\\ 
-- (.25-\epsilon/3) (\dot{b_l}/\omega-a_l) (p/\omega cos(\beta_w)\\ 
-+ q/\omega sin(\beta_w)) - .125 (-p/\omega sin(\beta_w)+q/\omega cos(\beta_w))^2 \\ 
--.125 (p/\omega cos(\beta_w)+q/\omega sin(\beta_w))^2 -  \\
-(.25-2/3 \epsilon+\epsilon^2/2) ((\dot{a_0}/\omega)^2 \\ 
-+ .5 ((\dot{a_l}/\omega+b_l)^2+(\dot{b_l}/\omega-a_l)^2))
-\label{torquecoeff}
-\end{multline}
-
-Please see \cite{chen1979simplified} for definition of terms.
-Although this model takes into account all of the design variables chosen for this project, it is difficult to optimize because of the complexity of the functions.
-\textit{Currently, this model fails to converge to a solution, so we are working with it to find the source of the issue.}
-
 \section{OPTIMIZATION}

### 2016-04-12 paper.tex
started reworking motivation and background

--- 2016-03-24 paper.tex
@@
 
+Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank respectively.
+The control of such a vehicle would also fundamentally change.
+Electric multirotors are controlled by spinning the rotors faster at slower to produce varying thrust and torque.
+Gad engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.
+
 Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
@@
 
-All commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
-Although these batteries are very good and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
-The average specific energy for Lithium-Polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
-This means that a gasoline-fueled multirotor can potential fly much longer than a battery powered multirotor because it can carry more energy onboard.
+Nearly all commercial and consumer multirotor UAVs are powered by lithium-polymer batteries because of their high energy capacity to weight ratios.
+Although these batteries are very useful and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
+The average specific energy for lithium-polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
+This means that a gasoline-fueled multirotor can potentially fly much longer than a battery powered vehicle because it can carry more energy.
 
+The gass powered quadcopter will have variable pitch propellers would be  
 \begin{itemize}
@@
 All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.
-We collected data about popular two stroke RC aircraft engines from several sources to produce the below plots.
+We collected data about popular two stroke RC aircraft engines from several sources to produce the models.
+One source was Desert Aircraft \cite{da2016}.
 
@@
 	\includegraphics[width=0.5\textwidth]{mass.png}
-	\caption{Stand in figure showing the data used to obtain the empirical based surrogate model of engine mass as a function of displacement size.}
+	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}}
 		\label{fig:mass}
@@
 	\includegraphics[width=0.5\textwidth]{max_power.png}
-	\caption{Stand in figure showing the data used to obtain the empirical based surrogate model of engine maximum power as a function of displacement size.}
+	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size. Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016}}
 		\label{fig:power}

### 2016-04-12 paper.tex

--- 2016-04-12 paper.tex
@@
 
-Nearly all commercial and consumer multirotor UAVs are powered by Lithium-Polymer batteries because of their high energy capacity to weight ratios.
-Although these batteries are very good and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
-The average specific energy for Lithium-Polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
-This means that a gasoline-fueled multirotor can potential fly much longer than a battery powered multirotor because it can carry more energy onboard.
+Nearly all commercial and consumer multirotor UAVs are powered by lithium-polymer batteries because of their high energy capacity to weight ratios.
+Although these batteries are very useful and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
+The average specific energy for lithium-polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
+This means that a gasoline-fueled multirotor can potentially fly much longer than a battery powered vehicle because it can carry more energy.
 
+The gass powered quadcopter will have variable pitch propellers would be  
 \begin{itemize}

### 2016-04-12 paper.tex
worked on Engine models, background, and various other little things

@@
 Electric multirotors are controlled by spinning the rotors faster at slower to produce varying thrust and torque.
-Gad engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.
+Gas engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.
 
@@
 
-There are many commercially available parts for a gas quad, including engines and variable pitch helicopter propellers.
+There are many commercially available parts for a gas quad, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
+However a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.
+
+FAA requirements...
 
@@
 Because the quad is going to be build using commercially available off-the-shelf parts the optimization must be closely based on reality.
-This means we only can pick the size of engine that we are going to use.
+The analysis must, therefore, either be based on observations from available hardware or attempt analyze the hardware in its available configuration.
+The following sections present some surrogate models for approximating engine parameters as well as aerodynamic analysis of common variable pitch RC propellers.
+
+\subsection{Engine}
+
+The designer essentially can only choose the size of engine.
 All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.
-We collected data about popular two stroke RC aircraft engines from several sources to produce the models.
+Data was collected for popular two stroke RC aircraft engines from several sources to produce the models for other unknown parameters.
 One source was Desert Aircraft \cite{da2016}.
+Figure~\ref{fig:mass} shows the collected data for engine size and mass.
+We performed a linear least squares bast fit to produce a model for use in our analysis.
 
-\begin{figurehere}
+\begin{figure}
 	\includegraphics[width=0.5\textwidth]{mass.png}
-	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}}
+	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}. $\ y = ax + b \quad a = 0.02145 \quad b = 0.25412$}
 		\label{fig:mass}
-\end{figurehere}
+\end{figure}
 
-\begin{figurehere}
+Data was also used to produce a model of the power output of the engine as a function of size.
+Figure~\ref{fig:power} shows that collected data for engine size and power.
+Again we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}.
+However both models were inadequate for our analysis because they contain a non-zero y-intercept.
+A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to fly infinitely.
+To fix this deficiency we applied a quadratic fit to the data and forced it to do through the origin.
+
+\begin{figure}
 	\includegraphics[width=0.5\textwidth]{max_power.png}
-	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size. Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016}}
+	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size.
+Notice the linear models have a non-zero y intercept but the quadratic fit goes through the origin.
+Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016} $\ y = ax^2 + bx \quad a = -0.04548 \quad b = 79.504$}
 		\label{fig:power}
-\end{figurehere}
+\end{figure}
 
-The engine efficiency (friction and heat transfer losses), however, is a function of power capability of the engine (max power) and the power engine is required to produce.
-For this we will need a better model.
-\textit{We are still working on this.
-Since efficiency numbers for RC engines aren't publicly available, we need a scientific model.
-For now we just assume a constant efficiency.
-We plan to work on this in the future.}
-
-Although the off-the-shelf constraint also applies to choosing the rotor parameters, continuous variables approximating the rotor specs are used and then analyzed as follows.
+The engine efficiency was not modeled for this project and was assumed a constant parameter.  Within the analysis the shaft output power of the engine was 15\% of the power consumed from the fuel.
 

### 2016-04-12 paper.tex
worked on the paper intro

@@
 %\usepackage{biber}
+
+%for citing webcites
+\usepackage{cite}
+\usepackage{url}
 
@@
 
-\begin{itemize}
-	\item{This paper looks at the why and how we are going to optimize a multirotor platform to maximize flight time.}
-	\item{Talk about how the variables that are going to be optimized are rotor radius, fuel consumption rate, number of blade per rotor, number of rotors, rotor speed, and rotor pitch.}
-	\item{Talk about why these variables were selected as design variables.}
-	\item{Talk about some physical constraints, especially those imposed by the FAA.}
-\end{itemize}
+The creation of a multirotor aircaft capable of both carrying a relatively large payload and flying for a long duration would be very desirable fo many applications.
+
+Designing a gas powered multirotor aircraft requires several design decisions.
+When considering the total flight time of a the aircraft, foremost of these design decisions are the choice of rotor, engine and gas tank size.
+These, and other, trade-offs are hard for a designer to simultaneously consider.
+
+This paper presents analysis and optimization methods for designing a gas powered multirotor aircraft.
+The analysis focuses on modeling the engine and aerodynamic components with a limited consideration of aircraft structure.
+These analyses are then used within multi-dimensional optimization to produce a vehicle design capable of long duration flight.
 
@@
 
+Multirotor aircraft are useful for a variety of applications.
+Because multirotors aircraft can fly vertically they do not required landing strips.
+They can carry sensor equipment for scientific or surveillance purposes.
+Multirotor use in package delivery is being studied by several large companies such as Google, Amazon and DHL \cite{Amazon2014}.
+
 Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank respectively.
 The control of such a vehicle would also fundamentally change.
-Electric multirotors are controlled by spinning the rotors faster at slower to produce varying thrust and torque.
-Gas engines are unable to rev up as fast as electric motors meaning a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.
+Electric multirotors are controlled by varying propeller rotational speed to produce varying thrust and torque.
+Gas engines lack the ability to vary speed quickly enough to control the attitude of a multirotor aircraft.
+This means a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.
 
@@
 There are many commercially available parts for a gas quad, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
-However a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.
+However, a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.
 
-FAA requirements...
+FAA requirements \cite{faa2016} specify that model aircraft must weigh less then 55 lbs (25 kg) or else they require special airworthiness certificates.
 

### 2016-04-12 paper.tex
Finished Aero, started conclusions

--- 2016-04-12 paper.tex
@@
 Combined differential blade element and momentum theory for nonuniform inflow was used to calculate the thrust and the torque differentially across each rotor blade \cite{bramwell2001bramwell}.
-The differential thrust is given by
+Using blade element theory, the differential thrust is given by
 
 \begin{equation}
-	dT = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr,
+	dT_{prod} = \frac{1}{2} \rho a \Omega^2 r^2 (\theta - \phi) c \  dr,
 	\label{thrust_eqn}
@@
 The total thrust on each rotor is calculated by integrating equation \ref{thrust_eqn} across the rotor radius, \textit{R}.
-The total thrust for the multirotor can then be calculated by multiplying the blade thrust by the number of \textit{N} rotors.
+The total thrust for the multirotor, $T_{prod}$ can then be calculated by multiplying the blade thrust by the number of \textit{N} rotors.
 
@@
 \begin{equation}
-	dQ = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr,
+	dQ_{req} = \frac{1}{2} \rho \Omega^2 r^3 c (\delta + \phi C_L) \ dr,
 	\label{torque_eqn}
@@
 
-where $\delta$ is the local blade drag coefficient, and $C_L$ is the lift coefficient. The total torque for each blade and the multirotor can be found using the same method for calculating the thrust.
+where $\delta$ is the local blade drag coefficient, and $C_L$ is the lift coefficient. The total torque for the multirotor , $Q_{req}$, can be found using the same method for calculating the thrust.
 
-The differential power produced by each rotor is given by
+The differential power produced by each rotor is then given by
 
 \begin{equation}
-	dP = dQ\Omega,
+	dP_{req} = dQ\Omega,
+	\label{power_eqn}
 \end{equation}
 
-and the total power can be calculated using the same method as thrust and torque.
+and the total power required by the rotors, $P_{req}$, can be calculated using the same method as thrust and torque.
+
+Equations \ref{thrust_eqn}, \ref{torque_eqn} and \ref{power_eqn} are used in the constraints, as discussed in the Optimization section.
+In practice, the thrust, torque, and power were integrated in a \textit{for loop} that divided each blade into 10 elements and then summed the differential values for each element.
+More divisions were made, but it greatly increased the computation time of the optimization algorithm and the results were approximately the same.
+
+Since the goal of this project is to optimize the flight time of a multirotor and then build it using existing remote controlled aircraft components, a NACA 0012 airfoil was used in the analysis for calculating the coefficient of lift, $C_L$, and the blade element drag coefficient, $\delta$.
+Data points for this airfoil were generated using Xfoil.
+This data was then fit with polynomials to approximately represent $C_L$ and $\delta$ across the range of the angle of attack, $\alpha$, continuously, where $\alpha = \theta - \phi$.
+
+Momentum theory allows for a relatively simple calculation of the inflow angle for each element of the blade.
+This calculation is an approximation, but it is sufficient for our lower fidelity model in designing the rotor system.
+The inflow angle is given by
+
+\begin{equation}
+\phi^2 = (\frac{a\sigma}{8})(\theta-\phi),
+\label{inflow_eqn}
+\end{equation}
+
+where
+
+$$\sigma = \frac{bc}{\pi r}$$
+
+is the blade solidity ratio based on local rotor radius, and \textit{b} is the number of blades in each ratio.
 
@@
 
-Our design variables are rotor radius, fuel consumption rate, number of blades per rotor, number of rotors, rotor speed, and rotor pitch.
+Our design variables are rotor radius, fuel consumption rate, number of blades per rotor, number of rotors, rotor speed, rotor pitch, engine size, and fuel capacity.
 Rotor radius and number of blades per rotor, and number of rotors were chosen because this is an easy parameter to vary on multirotors due to the availability of numerous rotor blades available for small rotor aircraft.
@@
 
-We have found that the flight time of a gas-quad can be maximized by constructing with the following design.
+We have found that if the design variables are left unbound and all of the constraints are still being met, the maximum flight time could be 11.32 hours.
+The values of these design variables can be seen in table \ref{table:unbound_results}.
 
@@
 \centering
-\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|C|}
+\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|}
 \hline
-       & displacement & fuel rate & rotor radius & other \\ \hline
-  & 100 & 30 & 250 & 150 \\ \hline
+       \# of Blades & Rotor Radius & Rotor Speed & \# of Rotors \\ \hline
+  100 & 30 & 250 & 150 \\ \hline
 \end{tabulary}
 \caption{Optimal design variables}
-\label{table:fw_loop_rates}
+\label{table:unbound_results}
 \end{tablehere}

### 2016-04-12 paper.tex
paper motivation and methods section

--- 2016-04-12 paper.tex
@@
 
-Currently, there are numerous multirotor UAVs that are available commercially and to consumers that are capable of lifting small payloads for an endurance of approximately 30 minutes.
+Currently, there are numerous multirotor UAVs that are available commercially and to consumers that are capable of lifting small payloads for an endurance of approximately 30 minutes or less.
 While these current multirotors are good for short endurance missions, there are few options for larger, higher payload capacity multirotors that are able to maintain flight for longer than an hour.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
+\section{BACKGROUND}
 
@@
 
+Variable pitch mulitrotor aircraft are not as common as variable speed multirotors but they have been explored to some degree.
 Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
-There is a small, electric variable pitch quad commercially available \cite{stingray2016}.
+There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and 3D maneuvers.
 There are also several hobbyist attempts at gas powered variable pitch quads \cite{diy2016}, \cite{hackaday2016}.
-
 In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload.
 
-There are many commercially available parts for a gas quad, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
+There are many commercially available parts that could be used to create a gas powered multirotor, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
 However, a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.
@@
 
-Nearly all commercial and consumer multirotor UAVs are powered by lithium-polymer batteries because of their high energy capacity to weight ratios.
+Nearly all commercial and consumer multirotor UAVs are powered by lithium-polymer batteries because of their relatively high energy capacity to weight ratios.
 Although these batteries are very useful and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
-The average specific energy for lithium-polymer batteries is up to 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
+The average specific energy for lithium-polymer batteries is approximately 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
 This means that a gasoline-fueled multirotor can potentially fly much longer than a battery powered vehicle because it can carry more energy.
-
-The gass powered quadcopter will have variable pitch propellers would be  
-\begin{itemize}
-	\item{Will talk more about some of the details using a gas engine entails in the paper.}
-	\item{Speak briefly on the reasons for requiring variable pitch with a gas engine.} 
-	\item{Speak on current platforms and capabilities and explain how this could fundamentally change the way multirotors are used in commercial applications, but not really for consumers.}
-\end{itemize}
+%Is there a refernce for these numbers???
 
@@
 	\begin{center}
-		\includegraphics[width=.40\textwidth]{current_capabilities.png}
-		\caption{\textit{Graphic showing current platforms and how our would be better.}}
-		\label{current_cap}
+		\includegraphics[width=.5\textwidth]{current_capabilities.png}
+		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project.}} %This data probably needs a reference !!!!
+		\label{fig:current_cap}
 	\end{center}
 \end{figurehere}
+
+The goal of this project is to expand the flight envelope of multirotor aircraft.
+Figure~\ref{fig:current_cap} demonstrates the current capabilities of consumer and commercial multirotors and depicted the goal of this project.
+We believe that a multirotor that could carry larger payloads for longer periods of time would greatly expand the type and number of applications for these vehicles.
+A multirotor with this capability would fundamentally change the way multirotors are used in commercial applications.
 
@@
 
-The engine efficiency was not modeled for this project and was assumed a constant parameter.  Within the analysis the shaft output power of the engine was 15\% of the power consumed from the fuel.
+The engine efficiency was not modeled for this project and was assumed as a constant parameter.
+Throughout the analysis, the shaft output power of the engine was 15\% of the power consumed from the fuel.
 

### 2016-04-13 paper.tex
Added results for matlab opt

--- 2016-04-12 paper.tex
@@
   \# of Blades & Rotor Radius & Rotor Speed & \# of Rotors \\ \hline
-  100 & 30 & 250 & 150 \\ \hline
+  3 & 2.306 & 39 & 8 \\ \hline
+  Blade Pitch & Fuel Rate & Engine Size & Fuel Capacity \\ \hline
+  22 & 2.01e-4 & 11.31 & 12.77 \\ \hline
 \end{tabulary}
-\caption{Optimal design variables}
+\caption{Optimal design variables for loosely bound constraints.
+Rotor radius is in meters, rotor speed is in RPM, blade pitch is in degrees, fuel rate is in kg/s, engine size is in cc, fuel capacity is in liters.}
 \label{table:unbound_results}

### 2016-04-13 paper.tex
finished a draft of the optimization section

@@
 
-$$\sigma = \frac{bc}{\pi r}$$
+\[ \sigma = \frac{bc}{\pi r} \]
 
@@
 
-Our design variables are rotor radius, fuel consumption rate, number of blades per rotor, number of rotors, rotor speed, rotor pitch, engine size, and fuel capacity.
-Rotor radius and number of blades per rotor, and number of rotors were chosen because this is an easy parameter to vary on multirotors due to the availability of numerous rotor blades available for small rotor aircraft.
-These variables also have a large influence on the flight time of the multirotor.
-Rotor speed and rotor pitch were chosen because these variables are easy to change on the physical system through selecting the proper gearing ratios.
-Fuel consumption rate seems like an odd variable for an optimizer because intuitively an optimizer would just minimize this value to maximize the flight time; however, fuel consumption rate is constrained by the size of engine required to produce the amount of power needed to provide enough lift for the multirotor.
-All of these design variables, however, have bound constraints that keep them in a reasonable space of commercially available parts/performance.
+Within the optimization there are two sets of design variables.
+First, the design variables that influence the size of the vehicle: the rotor radius $R_{rot}$, the engine size $S_{eng}$, and the fuel capacity $f_{capacity}$.
+Then, there are the design variables that are used to determine feasibility of the design: fuel burn rate $f_{rate}$, rotor pitch angle $\theta_{pitch}$, and propeller head speed $\omega_{rot}$.
+Number of blades per rotor and number of rotors could also be treated as design variables.
+However, since the optimizer will often tent toward more rotors and adding more rotors would have unmodeled adverse affects on cost, complexity, and weight these variables are fixed as parameters in the analysis.
+Moving forward we will assume four sets of two blade rotors.
 
-Setup the formal optimization problem
+%The flight time of the vehicle is only directly proportional to two of the design variables: the fuel capacity $C_{fuel}$ and the fuel burn rate $r_{fuel}$.
+Once the fuel is all burned then the vehicle must land.
+Through the constraints imposed on the optimization the flight time becomes indirectly a function of the rest of the design variables.
+
+We now setup the formal optimization problem
 \begin{equation}
-min. \quad f\of{x} = -t_{flight}\of{x}
+	\begin{split}
+		min. & \quad f\of{x} = -t_{flight}\of{x} \\
+		w.r.t. & \quad x = [R_{rot}, S_{eng}, f_{capacity}, f_{rate}, \theta_{pitch}, \omega_{rot}]^T \\
+		s.t. & \quad cons\of{x} \leq 0 
+	\end{split}
 	\label{eq:objective}
 \end{equation}
-\begin{equation}
-w.r.t. \quad x = [variables]^T
-\label{eq:vars}
-\end{equation}
-\begin{equation}
-s.t. \quad cons\of{x} \leq 0 
-\label{eq:constrants}
-\end{equation}
 
-Flight time was calculated by 
-\begin{equation}
-	t_{flight} = K_{energy}*f_{capacity}/f_{rate}.
-\end{equation}
+Where flight time was calculated by 
+\[
+	t_{flight} = K_{energy}\frac{f_{capacity}}{f_{rate}}.
+\]
 Without constraints, the solution to maximizing flight time would be trivial to calculate as well as impossible to construct.
-Therefore to optimize the system we must use constraints that ensure the quad is feasible.
-The constrains are on the power produced by the engine ($P_{prod}$), power required by the prop ($P_{req}$), thrust produced by the prop ($T_{prod}$), lift required by the quad ($T_{req}$) and maximum power capabilities of the engine ($P_{max}$).
-Where $P_{preq} \leq P_{prod}$, $T_{req} \leq T_{prod}$, and $P_{prod} \leq P_{max}$.
+Therefore to optimize the system we must use constraints that ensure the quad is feasible and legal, according to the FAA \cite{faa2016}.
+The constrains are on the power produced by the engine ($P_{shaft}$), power required by the prop ($P_{rot}$), thrust produced by the prop ($T_{prod}$), total weight of the vehicle ($W_{total}$) and maximum power capabilities of the engine ($P_{max}$).
+Where the following relationships hold
 
-We then will discuss the following:
-\begin{itemize}
-	\item{scaling of the design variables}
-	\item{how we got gradients}
-	\item{the optimization method that we utilized}
-	\item{python vs matlab benefits and challenges}
-\end{itemize}
+\[
+	\begin{split}
+	P_{rot} & \leq P_{shaft} \\
+	W_{total} & \leq T_{prod} \\ 
+	P_{shaft} & \leq P_{max} \\
+	W_{total} & \leq 55 lbs
+	\end{split}
+\]
 
-Since we can only optimize for maximum flight time give a payload (parameter) we can further explore how the characteristics of a gas quad change as we optimize the flight time for different payloads.
-It may turn out that the optimal quad parameters change greatly as it is optimized for different payload.
-In this case we could choose several payloads and optimize over all of then together.
-\textit{We will present these results with a figure and some discussion.}
+The above aerodynamic model is used to calculate $P_{rot}$ and $T_{prod}$.
+$P_{shaft}$ is calculated from the fuel burn rate and engine efficiency.
+$W_{total}$ is the combined weight of the engine, vehicle frame, fuel, and payload.
+$P_max$ comes from the surrogate engine power model depicted in figure~\ref{fig:power}.
 
-\begin{figurehere}
+The design variables were scaled for input in to the optimizer.
+The optimizers we used were the matlab's fmincon function using interior-point algorithm and the python pyoptsparse package using SNOPT algorithum.
+Gradients were found using auto-differentiation in Matlab and finite differencing in python.
+
+Our optimization found the maximum flight time given a payload, but another formulation would be to find the maximum payload given a specified flight time.
+This means we could create a Pareto frontier of this multi-objective optimization.
+However, to simplify we can create figure~\ref{fig:payload} by varying the payload parameter within our optimization (not a true Pareto frontier because payload was not specifically optimized).
+We then can further explore how the characteristics of a gas quad change as we optimize the flight time for different payloads.
+For example, figure~\ref{fig:eng_v_pl} shows how the optimal engine size changes as the payload parameter varies.
+
+\begin{figure}
 	\includegraphics[width=0.5\textwidth]{pareto_front.png}
-	\caption{Stand in figure showing optimal flight time as a function of payload.}
+	\caption{Pareto fronter like curve found by optimizing for flight time given specified payloads.}
 		\label{fig:payload}
-\end{figurehere}
+\end{figure}
 
-%Here we can do some discussion on constraint sensitivity - aerodynamics, engine efficiencies...
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{engine_size_vs_payload.png}
+	\caption{Optimal engine size (cc) given a specified payload.}
+		\label{fig:eng_v_pl}
+\end{figure}
+
+Because our final design requires a single optimal solution and not a range of solutions that depend on our payload capacity we then simultaneously optimize over several payloads and optimize over all of then together.
 

### 2016-04-13 paper.tex
Edits through aerodynamics

--- 2016-04-13 paper.tex
@@
 
-Currently, there are numerous multirotor UAVs that are available commercially and to consumers that are capable of lifting small payloads for an endurance of approximately 30 minutes or less.
+Currently, there are numerous multirotor unmanned aerial vehicles (UAVs) that are available to commercial operators and to consumers that are capable of lifting small payloads for an endurance of approximately 30 minutes or less.
 While these current multirotors are good for short endurance missions, there are few options for larger, higher payload capacity multirotors that are able to maintain flight for longer than an hour.
@@
 
-The creation of a multirotor aircaft capable of both carrying a relatively large payload and flying for a long duration would be very desirable fo many applications.
+The creation of a multirotor aircaft capable of both carrying a relatively large payload and flying for a long duration would be very desirable for many applications.
 
 Designing a gas powered multirotor aircraft requires several design decisions.
-When considering the total flight time of a the aircraft, foremost of these design decisions are the choice of rotor, engine and gas tank size.
-These, and other, trade-offs are hard for a designer to simultaneously consider.
+When considering the total flight time of a the aircraft, foremost of these design decisions are the choices of rotor, engine and gas tank size.
+These design decisions along with several other aircraft parameters all have trade-offs that makes it difficult for a designer to simultaneously consider which trade-offs will allow for the longest flight time and/or greatest payload capacity.
 
@@
 Multirotor aircraft are useful for a variety of applications.
-Because multirotors aircraft can fly vertically they do not required landing strips.
+Because multirotors can fly vertically, they do not require landing strips.
 They can carry sensor equipment for scientific or surveillance purposes.
@@
 
-Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank respectively.
+Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank, respectively.
 The control of such a vehicle would also fundamentally change.
-Electric multirotors are controlled by varying propeller rotational speed to produce varying thrust and torque.
+Electric multirotors are controlled by varying propeller rotational speed to produce varying thrust and torque, which in turn controls the attitude of the aircraft.
 Gas engines lack the ability to vary speed quickly enough to control the attitude of a multirotor aircraft.
@@
 Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
-There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and 3D maneuvers.
-There are also several hobbyist attempts at gas powered variable pitch quads \cite{diy2016}, \cite{hackaday2016}.
-In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload.
+There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and inverted maneuvers.
+There are also several hobbyist attempts at gas powered variable pitch quad-rotor (quads) \cite{diy2016}, \cite{hackaday2016}.
+In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload for commercial applications.
 
-There are many commercially available parts that could be used to create a gas powered multirotor, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
+There are many commercially available components that could be used to create a gas powered multirotor, including engines \cite{da2016} and variable pitch helicopter propellers \cite{align2016}.
 However, a gas powered multirotor would required a custom frame for mounting the engine and supplying power from the engine to the individual rotors.
 
-FAA requirements \cite{faa2016} specify that model aircraft must weigh less then 55 lbs (25 kg) or else they require special airworthiness certificates.
+FAA requirements \cite{faa2016} specify that model aircraft and UAVs of any sort must weigh less then 55 lbs (25 kg).
+UAVs or model aircraft weighing more than 55 lbs require special airworthiness certificates and are much more heavily regulated.
 
@@
 Nearly all commercial and consumer multirotor UAVs are powered by lithium-polymer batteries because of their relatively high energy capacity to weight ratios.
-Although these batteries are very useful and reliable, they still offer much lower specific energy ratios then that of liquid fuels such as gasoline.
-The average specific energy for lithium-polymer batteries is approximately 0.95 MJ/kg, where gasoline is 46.4 MJ/kg.
+Although these batteries are very useful and reliable, they still offer much lower specific energy ratios than that of liquid fuels such as gasoline.
+The average specific energy for lithium-polymer batteries is approximately 0.95 MJ/kg \cite{Battuni2016}, where gasoline is 46.4 MJ/kg \cite{energyed2016}.
 This means that a gasoline-fueled multirotor can potentially fly much longer than a battery powered vehicle because it can carry more energy.
-%Is there a refernce for these numbers???
 
@@
 		\includegraphics[width=.5\textwidth]{current_capabilities.png}
-		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project.}} %This data probably needs a reference !!!!
+		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project.
+The data points shown are specifications given by manufacturers of popular commercial and consumer UAVs and is collected from \cite{drones2016}.}}
 		\label{fig:current_cap}
@@
 The goal of this project is to expand the flight envelope of multirotor aircraft.
-Figure~\ref{fig:current_cap} demonstrates the current capabilities of consumer and commercial multirotors and depicted the goal of this project.
+Figure~\ref{fig:current_cap} demonstrates the current capabilities of consumer and commercial multirotors and depicts the goal of this project.
 We believe that a multirotor that could carry larger payloads for longer periods of time would greatly expand the type and number of applications for these vehicles.
@@
 
-Because the quad is going to be build using commercially available off-the-shelf parts the optimization must be closely based on reality.
-The analysis must, therefore, either be based on observations from available hardware or attempt analyze the hardware in its available configuration.
-The following sections present some surrogate models for approximating engine parameters as well as aerodynamic analysis of common variable pitch RC propellers.
+Because the quad is going to be built using commercially available off-the-shelf components, the optimization must be closely based on a realistic model.
+Therefore, the analysis must either be based on observations from available hardware or attempt to analyze the hardware in its available configurations.
+The following sections present some surrogate models for approximating engine parameters as well as an aerodynamic analysis of common variable pitch RC propellers.
 
@@
 
-The designer essentially can only choose the size of engine.
-All the other data about the engine (mass, maximum power, etc.) have to be a function of displacement size.
+The designer essentially can only choose the size of engine when choosing a power plant.
+All the other data about the engine (mass, maximum power, etc.) have to be functions of displacement size.
 Data was collected for popular two stroke RC aircraft engines from several sources to produce the models for other unknown parameters.
-One source was Desert Aircraft \cite{da2016}.
+One source for engine data was Desert Aircraft \cite{da2016}.
 Figure~\ref{fig:mass} shows the collected data for engine size and mass.
-We performed a linear least squares bast fit to produce a model for use in our analysis.
+We performed a linear least squares best fit to produce a model for use in our analysis.
 
@@
 	\includegraphics[width=0.5\textwidth]{mass.png}
-	\caption{The empirical based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}. $\ y = ax + b \quad a = 0.02145 \quad b = 0.25412$}
+	\caption{The empirical data based surrogate model of engine mass as a function of displacement size. Data from desertaircraft.com \cite{da2016}. $\ y = ax + b \quad a = 0.02145 \quad b = 0.25412$}
 		\label{fig:mass}
@@
 Data was also used to produce a model of the power output of the engine as a function of size.
-Figure~\ref{fig:power} shows that collected data for engine size and power.
-Again we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}.
-However both models were inadequate for our analysis because they contain a non-zero y-intercept.
+Figure~\ref{fig:power} shows the collected data for engine size and power.
+Again, we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}.
+However, both models were inadequate for our analysis because they contain a non-zero y-intercept.
 A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to fly infinitely.
-To fix this deficiency we applied a quadratic fit to the data and forced it to do through the origin.
+To correct for this issue, we applied a quadratic fit to the data and forced it to intercept through the origin.
 
@@
 	\caption{Empirical based surrogate model of engine maximum power as a function of displacement size.
-Notice the linear models have a non-zero y intercept but the quadratic fit goes through the origin.
+Notice the linear models have a non-zero y-intercept but the quadratic fit goes through the origin.
 Data from desertaircraft.com \cite{da2016} and barnardmicrosystems.com \cite{barnardmiro2016} $\ y = ax^2 + bx \quad a = -0.04548 \quad b = 79.504$}
@@
 Many of the existing aerodynamic models used in the aerospace industry to represent helicopter rotors are extremely complex and require specialized software tools (many of which are proprietary).
-High fidelity aerodynamic models generally require lots of computational power; so in order to focus on optimization methods rather than the complex aerodynamics of the system, several simplifications were made.
+High fidelity aerodynamic models generally require lots of computational power; so in order to focus on optimization methods, rather than the complex aerodynamics of the system, several simplifications were made in analyzing the inflow angle and the down-wind air interactions of the propeller.
+Also, the model used assumes the multirotor is in a hover, since that seems to be the primary use of multirotor aircraft.
 

### 2016-04-13 paper.tex
added draft of results

@@
 
-Because our final design requires a single optimal solution and not a range of solutions that depend on our payload capacity we then simultaneously optimize over several payloads and optimize over all of then together.
+Because the final design requires a single optimal solution and not a range of solutions that depend on our payload capacity we then simultaneously optimize at several payloads.
+This means the optimization must maximize the total combined flight time while stratifying constraints for operating points.
 
@@
 
-We have found that if the design variables are left unbound and all of the constraints are still being met, the maximum flight time could be 11.32 hours.
-The values of these design variables can be seen in table \ref{table:unbound_results}.
+When simultaneously optimizing with 3, 5, and 7 kg payloads we get an optimal solution.
+The solution flight time is at least 42.46 minutes with a payload less then 7 kg.
+The values of the optimal design variables can be seen in table~\ref{table:unbound_results}.
 
@@
 \centering
-\begin{tabulary}{0.5\textwidth}{|C|C|C|C|C|}
+\vspace{5mm}
+	\begin{tabulary}{0.5\textwidth}{|C|C|}
 		\hline
-  \# of Blades & Rotor Radius & Rotor Speed & \# of Rotors \\ \hline
-  3 & 2.306 & 39 & 8 \\ \hline
-  Blade Pitch & Fuel Rate & Engine Size & Fuel Capacity \\ \hline
-  22 & 2.01e-4 & 11.31 & 12.77 \\ \hline
+		Design Variable & Optimal Value \\ \hline \hline
+		Rotor radius (m) & 2.156 \\ \hline
+		Engine size (cc) & 176.9 \\ \hline
+		Fuel Capacity (kg) & 6.45 \\ \hline
+		Fuel burn rate (g/s) & 1.898 \\ \hline
+		Rotor pitch (deg) & 17.6 \\ \hline
+		Head speed (rpm) & 135.1 \\ \hline
 	\end{tabulary}
-\caption{Optimal design variables for loosely bound constraints.
-Rotor radius is in meters, rotor speed is in RPM, blade pitch is in degrees, fuel rate is in kg/s, engine size is in cc, fuel capacity is in liters.}
+\caption{Optimal design variables for loosely bound constraints.}
 \label{table:unbound_results}
 \end{tablehere}
+
+The above solution is a vehicle with an unreasonably large slow moving propellers.
+While bounding our rotor size would give a more physically manageable solution, it would have also meant our optimization was largely an aerodynamic analysis with a small optimization of engine size.
+Leaving the variable loosely bounded gives more insight about the trade offs of the vehicle makes for efficiency.
+
+We also found that the solutions varied greatly with very small changes in our aerodynamic analysis.
+Personal experience with RC aircraft lead us to believe that the head speed and pitch angles are too extreme.
+Further validation of our analysis would be required before a design such as this to be built.
 

### 2016-04-13 paper.tex
Edits through optimization

--- 2016-04-13 paper.tex
@@
 Many of the existing aerodynamic models used in the aerospace industry to represent helicopter rotors are extremely complex and require specialized software tools (many of which are proprietary).
-High fidelity aerodynamic models generally require lots of computational power; so in order to focus on optimization methods, rather than the complex aerodynamics of the system, several simplifications were made in analyzing the inflow angle and the down-wind air interactions of the propeller.
-Also, the model used assumes the multirotor is in a hover, since that seems to be the primary use of multirotor aircraft.
+High fidelity aerodynamic models generally require lots of computational power; so in order to focus on optimization methods, rather than the complex aerodynamic interactions of the system, several simplifications were made in analyzing the inflow angle and the down-wind air interactions of the propeller.
+Also, the model used assumes the multirotor is in hover, since that seems to be the primary use condition of multirotor aircraft.
 
@@
 The total thrust on each rotor is calculated by integrating equation \ref{thrust_eqn} across the rotor radius, \textit{R}.
-The total thrust for the multirotor, $T_{prod}$ can then be calculated by multiplying the blade thrust by the number of \textit{N} rotors.
+The total thrust for the multirotor, $T_{prod}$ can then be calculated by multiplying the rotor thrust by the number of \textit{N} rotors.
 
@@
 
-where $\delta$ is the local blade drag coefficient, and $C_L$ is the lift coefficient. The total torque for the multirotor , $Q_{req}$, can be found using the same method for calculating the thrust.
+where $\delta$ is the local blade drag coefficient, and $C_L$ is the lift coefficient.
+The total torque for the multirotor , $Q_{req}$, can be found using the same method for calculating the total thrust.
 
@@
 
-and the total power required by the rotors, $P_{req}$, can be calculated using the same method as thrust and torque.
+and the total power required by the rotors, $P_{req}$, can be calculated using the same method as total thrust and torque \cite{venkatesan2014fundamentals}.
 
@@
 In practice, the thrust, torque, and power were integrated in a \textit{for loop} that divided each blade into 10 elements and then summed the differential values for each element.
-More divisions were made, but it greatly increased the computation time of the optimization algorithm and the results were approximately the same.
+More divisions were made initially, but it greatly increased the computation time of the optimization algorithm and the results were approximately the same.
 
 Since the goal of this project is to optimize the flight time of a multirotor and then build it using existing remote controlled aircraft components, a NACA 0012 airfoil was used in the analysis for calculating the coefficient of lift, $C_L$, and the blade element drag coefficient, $\delta$.
-Data points for this airfoil were generated using Xfoil.
-This data was then fit with polynomials to approximately represent $C_L$ and $\delta$ across the range of the angle of attack, $\alpha$, continuously, where $\alpha = \theta - \phi$.
+This airfoil approximately represents the airfoil of many commonly used remote control helicopters \cite{mettler2013identification}.
+Data points for this airfoil were generated using Xfoil, an aerodynamics tool developed by The Massachusetts Institute of Technology.
+This data was then fit with polynomials to approximately represent $C_L$ and $\delta$ across a range of the angles of attack, $\alpha$, continuously, where $\alpha = \theta - \phi$.
 
 Momentum theory allows for a relatively simple calculation of the inflow angle for each element of the blade.
-This calculation is an approximation, but it is sufficient for our lower fidelity model in designing the rotor system.
+This calculation is an approximation, but it is sufficient for our low-fidelity model in designing the rotor system.
 The inflow angle is given by
@@
 
-is the blade solidity ratio based on local rotor radius, and \textit{b} is the number of blades in each ratio.
+is the blade solidity ratio based on local rotor radius, and \textit{b} is the number of blades in each rotor.
 
@@
 Then, there are the design variables that are used to determine feasibility of the design: fuel burn rate $f_{rate}$, rotor pitch angle $\theta_{pitch}$, and propeller head speed $\omega_{rot}$.
-Number of blades per rotor and number of rotors could also be treated as design variables.
-However, since the optimizer will often tent toward more rotors and adding more rotors would have unmodeled adverse affects on cost, complexity, and weight these variables are fixed as parameters in the analysis.
-Moving forward we will assume four sets of two blade rotors.
+Number of blades per rotor and number of rotors were also initially treated as design variables.
+However, with the current aerodynamic model, the optimizer will create as many rotors with as many blades as possible to minimize the rotational speed required of each rotor to reduce the overall drag.
+Since adding more rotors would have un-modeled adverse affects on cost, complexity, and weight these variables are fixed as parameters in the analysis.
+Moving forward we will assume four sets of two blade rotors because this is a common configuration that would be reasonably simple to build.
 
@@
 
-Where flight time was calculated by 
+where flight time was calculated by 
 \[
@@
 Therefore to optimize the system we must use constraints that ensure the quad is feasible and legal, according to the FAA \cite{faa2016}.
-The constrains are on the power produced by the engine ($P_{shaft}$), power required by the prop ($P_{rot}$), thrust produced by the prop ($T_{prod}$), total weight of the vehicle ($W_{total}$) and maximum power capabilities of the engine ($P_{max}$).
-Where the following relationships hold
+The constraints are on the power produced by the engine ($P_{shaft}$), power required by the rotors ($P_{rot}$), thrust produced by the rotors ($T_{prod}$), total weight of the vehicle ($W_{total}$) and maximum power capabilities of the engine ($P_{max}$).
+To use the constraints correctly, the following relationships must hold:
 
@@
 The design variables were scaled for input in to the optimizer.
-The optimizers we used were the matlab's fmincon function using interior-point algorithm and the python pyoptsparse package using SNOPT algorithum.
-Gradients were found using auto-differentiation in Matlab and finite differencing in python.
+The optimizers we used were the Matlab's \textit{fmincon} function using interior-point algorithm and the Python \textit{pyoptsparse} package using SNOPT algorithum.
+Gradients were found using automatic-differentiation in Matlab and finite-differencing in Python.
 
-Our optimization found the maximum flight time given a payload, but another formulation would be to find the maximum payload given a specified flight time.
+Our optimization algorithm found the maximum flight time given a payload, but another formulation would be to find the maximum payload given a specified flight time.
 This means we could create a Pareto frontier of this multi-objective optimization.

### 2016-04-13 paper.tex
added some conclustions

@@
 
-When simultaneously optimizing with 3, 5, and 7 kg payloads we get an optimal solution.
-The solution flight time is at least 42.46 minutes with a payload less then 7 kg.
+When simultaneously optimizing with 3, 5, and 7~kg payloads we get an optimal solution.
+The solution flight time is at least 42.46~minutes with a payload not greater then 7~kg.
 The values of the optimal design variables can be seen in table~\ref{table:unbound_results}.
@@
 
-lots of really good conclusions
+We have demonstrated a optimization of a gas powered variable pitch multirotor aircraft capable of carrying relatively large payloads for long durations.
+The optimization produced a design that is feasible according to our analysis.
+The design can carry up to 7~kg for at least 42.5~minutes.
+This result is far better then the current capability of vehicles on the market.
+
+The optimization analysis makes several assumptions and approximations that would need to be relaxed in order to better trust the solution.
+Further the optimization chose a design from continuous variables when only discrete off-the-shelf choices are available.
+
+Future work would include using better aerodynamic modeling and validation with testing, adding structural analysis to find the weight of the frame, and optimizing under uncertainty due to inaccuracies in analysis.
 

### 2016-04-13 paper.tex
Fixed refrence url's

--- 2016-04-13 paper.tex
@@
 
-Since the goal of this project is to optimize the flight time of a multirotor and then build it using existing remote controlled aircraft components, a NACA 0012 airfoil was used in the analysis for calculating the coefficient of lift, $C_L$, and the blade element drag coefficient, $\delta$.
+Since the goal of this project is to optimize the flight time of a multirotor and then build it using existing remote controlled aircraft components, a NACA 0012 airfoil (see \cite{abbott1959theory}) was used in the analysis for calculating the coefficient of lift, $C_L$, and the blade element drag coefficient, $\delta$.
 This airfoil approximately represents the airfoil of many commonly used remote control helicopters \cite{mettler2013identification}.
@@
 	\includegraphics[width=0.5\textwidth]{pareto_front.png}
-	\caption{Pareto fronter like curve found by optimizing for flight time given specified payloads.}
+	\caption{Pareto frontier-like curve found by optimizing for flight time given specified payloads.}
 		\label{fig:payload}
@@
 
-Because the final design requires a single optimal solution and not a range of solutions that depend on our payload capacity we then simultaneously optimize at several payloads.
-This means the optimization must maximize the total combined flight time while stratifying constraints for operating points.
+Because the final design requires a single optimal solution and not a range of solutions that depend on our payload capacity, we then simultaneously optimized at several payloads.
+This means the optimization must maximize the total combined flight time while satisfying constraints for operating points.
 
@@
 
-When simultaneously optimizing with 3, 5, and 7 kg payloads we get an optimal solution.
+When simultaneously optimizing with 3, 5, and 7 kg payloads our optimizer converges to an optimal solution.
 The solution flight time is at least 42.46 minutes with a payload less then 7 kg.
@@
 
-The above solution is a vehicle with an unreasonably large slow moving propellers.
+The above solution is a vehicle with unreasonably large, slow moving propellers.
 While bounding our rotor size would give a more physically manageable solution, it would have also meant our optimization was largely an aerodynamic analysis with a small optimization of engine size.
-Leaving the variable loosely bounded gives more insight about the trade offs of the vehicle makes for efficiency.
+Leaving the variable loosely bounded gives more insight about the trade-offs the vehicle makes for efficiency.
 
 We also found that the solutions varied greatly with very small changes in our aerodynamic analysis.
-Personal experience with RC aircraft lead us to believe that the head speed and pitch angles are too extreme.
-Further validation of our analysis would be required before a design such as this to be built.
+Personal experience with RC aircraft leads us to believe that the head speed and pitch angles are too extreme.
+Further validation of our analysis would be required before a design such as this were to be built.
 

### 2016-04-13 paper.tex

@@
 
-When simultaneously optimizing with 3, 5, and 7 kg payloads our optimizer converges to an optimal solution.
-The solution flight time is at least 42.46 minutes with a payload less then 7 kg.
+When simultaneously optimizing with 3, 5, and 7~kg payloads our optimizer converges to an optimal solution.
+The solution flight time is at least 42.46~minutes with a payload not greater than 7~kg.
 The values of the optimal design variables can be seen in table~\ref{table:unbound_results}.

### 2016-04-13 paper.tex
Edited conclusions

@@
 
-We have demonstrated a optimization of a gas powered variable pitch multirotor aircraft capable of carrying relatively large payloads for long durations.
+We have demonstrated an optimization of a gas powered variable pitch multirotor aircraft capable of carrying relatively large payloads for long durations.
 The optimization produced a design that is feasible according to our analysis.
 The design can carry up to 7~kg for at least 42.5~minutes.
-This result is far better then the current capability of vehicles on the market.
+This result is far better then the current capability of battery powered vehicles on the market.
 
-The optimization analysis makes several assumptions and approximations that would need to be relaxed in order to better trust the solution.
-Further the optimization chose a design from continuous variables when only discrete off-the-shelf choices are available.
+The optimization analysis makes several assumptions and approximations that would need to be improved in order to better trust the solution.
+Furthermore, the optimization chose a design from continuous variables when only discrete off-the-shelf choices are available.
 

### 2016-04-13 paper.tex
final small revisions

@@
 When considering the total flight time of a the aircraft, foremost of these design decisions are the choices of rotor, engine and gas tank size.
-These design decisions along with several other aircraft parameters all have trade-offs that makes it difficult for a designer to simultaneously consider which trade-offs will allow for the longest flight time and/or greatest payload capacity.
+These design decisions along with several other aircraft parameters all have trade-offs that makes it difficult for a designer to simultaneously consider which design choice will allow for the longest flight time and/or greatest payload capacity.
 
@@
 The analysis focuses on modeling the engine and aerodynamic components with a limited consideration of aircraft structure.
-These analyses are then used within multi-dimensional optimization to produce a vehicle design capable of long duration flight.
+These analyses are then used within multi-dimensional optimization to produce a vehicle design capable of both long duration flight and lifting heavy payloads.
 
@@
 Multirotor aircraft are useful for a variety of applications.
-Because multirotors can fly vertically, they do not require landing strips.
-They can carry sensor equipment for scientific or surveillance purposes.
-Multirotor use in package delivery is being studied by several large companies such as Google, Amazon and DHL \cite{Amazon2014}.
+This is because multirotors can hover, fly vertically, and fly horizontally.
+Further, they do not require landing strips and can carry sensor equipment for scientific or surveillance purposes.
+Multirotor use in package delivery is being considered by several large companies such as Google, Amazon and DHL \cite{Amazon2014}.
 
-Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank, respectively.
+Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank.
 The control of such a vehicle would also fundamentally change.
@@
 Variable pitch mulitrotor aircraft are not as common as variable speed multirotors but they have been explored to some degree.
-Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}.
+Control of variable pitch multirotor aircraft was studied by Cutler, et al.
+\cite{Cutler2012}\cite{Cutler2012a} and Segui-Gasco, et al.
+\cite{Segui-Gasco2014}.
 There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and inverted maneuvers.
-There are also several hobbyist attempts at gas powered variable pitch quad-rotor (quads) \cite{diy2016}, \cite{hackaday2016}.
-In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload for commercial applications.
+There are also several hobbyist attempts at gas powered variable pitch quad-rotors (quads) \cite{diy2016}, \cite{hackaday2016}.
+In our knowledge there has been no attempt to use a gas powered variable pitch multirotor to maximize the flight time for carrying a relatively large payload for commercial applications.
 
@@
 
-FAA requirements \cite{faa2016} specify that model aircraft and UAVs of any sort must weigh less then 55 lbs (25 kg).
+Federal Aviation Administration (FAA) requirements \cite{faa2016} specify that model aircraft and UAVs of any sort must weigh less then 55 lbs (25 kg).
 UAVs or model aircraft weighing more than 55 lbs require special airworthiness certificates and are much more heavily regulated.
@@
 \section{MOTIVATION}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=.5\textwidth]{current_capabilities.png}
+		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project.
+The data points shown are specifications given by manufacturers of popular commercial and consumer UAVs and is collected from \cite{drones2016}.}}
+		\label{fig:current_cap}
+	\end{center}
+\end{figure}
 
@@
 
-\begin{figurehere}
-	\begin{center}
-		\includegraphics[width=.5\textwidth]{current_capabilities.png}
-		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project.
-The data points shown are specifications given by manufacturers of popular commercial and consumer UAVs and is collected from \cite{drones2016}.}}
-		\label{fig:current_cap}
-	\end{center}
-\end{figurehere}
-
 The goal of this project is to expand the flight envelope of multirotor aircraft.
@@
 We believe that a multirotor that could carry larger payloads for longer periods of time would greatly expand the type and number of applications for these vehicles.
-A multirotor with this capability would fundamentally change the way multirotors are used in commercial applications.
+A multirotor with this capability would also fundamentally change the way multirotors are used in commercial applications.
 
@@
 However, both models were inadequate for our analysis because they contain a non-zero y-intercept.
-A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to fly infinitely.
+A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to have infinite flight time.
 To correct for this issue, we applied a quadratic fit to the data and forced it to intercept through the origin.
@@
 Moving forward we will assume four sets of two blade rotors because this is a common configuration that would be reasonably simple to build.
-
-%The flight time of the vehicle is only directly proportional to two of the design variables: the fuel capacity $C_{fuel}$ and the fuel burn rate $r_{fuel}$.
-Once the fuel is all burned then the vehicle must land.
-Through the constraints imposed on the optimization the flight time becomes indirectly a function of the rest of the design variables.
+Finally, although the selection of off-the-shelf components is a fundamentally discrete choice, for simplification within the optimization each design variable is treated as a continuous variable.
 
@@
 The optimization analysis makes several assumptions and approximations that would need to be improved in order to better trust the solution.
-Furthermore, the optimization chose a design from continuous variables when only discrete off-the-shelf choices are available.
-
-Future work would include using better aerodynamic modeling and validation with testing, adding structural analysis to find the weight of the frame, and optimizing under uncertainty due to inaccuracies in analysis.
+Furthermore, the optimization chose a design from a continuous space when only discrete off-the-shelf choices are available.
+Further selection of the components would be needed to build a quad through a discrete optimization algorithm or analyzing all designs close to the continuous optimum.
+Future work would also include using better aerodynamic modeling and validation with testing, adding structural analysis to find the weight of the frame, and optimizing under uncertainty due to inaccuracies in analysis.
 
